### 2022-02-22 README.Rmd
improved ggplot2 scales

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# Colors for all!

The **cols4all** is an R package in development for selecting color palettes. "Color for all" refers to our mission that colors should be usable for not just people with normal color vision, but also for people with color vision deficiency. This package contains palettes from several popular color palette series, such as Color Brewer, Viridis, and Kovesi, but also the lesser known [palettes from Paul Tol](https://personal.sron.nl/~pault/). Own palettes series can be added as well. Color palettes are well organized. Currently we support three types: *categorical* (qualitative) palettes, *sequential* palettes, and *diverging* palettes. In the near future, more palette types will be added, namely *cyclic*, *bivariate*, and *hierarchical*.

## Installation

**cols4all** will be available on CRAN soon! Until then it can be installed using:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("mtennekes/cols4all")
```

```{r}
getwd()
devtools::load_all(".")
```

## Get started

Navigation through the color palettes is easy:

```{r, eval=FALSE}
library(cols4all)
c4a_gui()
```

![categorical palettes)](https://user-images.githubusercontent.com/2444081/155185594-a52c361f-7113-40fe-82ea-94e782c27be4.png)
![sequential palettes)](https://user-images.githubusercontent.com/2444081/155185615-9356443d-8a18-40a6-bc9e-d2e0d8e26eb4.png)
![diverging palettes)](https://user-images.githubusercontent.com/2444081/155185638-3d05a045-c794-4eab-b994-88c0b58196e2.png)

Selecting a palette is easy. Palette names are in the format <series>.<palette_name>. So all letters are in small caps, and underscores are used to separate words.

```{r}
# Select the palette "kelly" with 7 colors
c4a("kelly", 7)

# find names of hcl palettes that are diverging
c4a_palettes(type = "div", series = "hcl")

# select purple green palette from the hcl series:
c4a("hcl.purple_green", 11)

# get the associated color for missing values
c4a_na("hcl.purple_green")
```

## Related R packages

The foundation of this package is another R package: [**colorspace**](https://colorspace.r-forge.r-project.org/). We use this package to analyse colors. For this purpose, we also use [**colorblindcheck**](https://github.com/Nowosad/colorblindcheck).

There are a few features that distinguishes **cols4all** from other (great) color palette packages, in particular [**pals**](https://kwstat.github.io/pals/) and [**paletteer**](https://github.com/EmilHvitfeldt/paletteer):

* Colors for missing values are made explicit. For this, either a greyscale color that is already contained in the palette is used, or a greyscale color is selected that is distinguisable, also for people with color vision deficiency.

* Palettes are made consistent with each other. For instance, sequential palettes (except for spectral/rainbow palettes) all start with the lightest color and end with the darkest color. Furthermore, black and white are removed from most categorical palettes for practical reasons: white or light grey is almost always used as background color and black for annotation.

* With the gui (see below), a user is able to analyse characteristics of color palettes, such as color-blind-friendliness, harmony, and the presence of intense colors.

* It will be possible to submit an own series of color palettes.

There will be native support for **ggplot** and **tmap** (as of the upcoming version 4)

### 2022-02-22 R/c4a_show.R

@@ -91,7 +91,7 @@ table_columns = function(type, show.scores) {
 #' @export
 #' @rdname c4a_gui
 #' @name c4a_gui
-c4a_show = function(type = c("cat", "seq", "div"), n = NULL, cvd.sim = c("none", "deutan", "protan", "tritan"), sort = "name", text.col = "same", series = NULL, contrast = NULL, include.na = TRUE, show.scores = FALSE, columns = NA) {
+c4a_show = function(type = c("cat", "seq", "div"), n = NULL, cvd.sim = c("none", "deutan", "protan", "tritan"), sort = "name", text.col = "same", series = NULL, contrast = NULL, include.na = FALSE, show.scores = FALSE, columns = NA) {
 
 	type = match.arg(type)
 	show.ranking = (!is.null(n))

### 2022-02-22 R/ggplot2.R

@@ -6,26 +6,86 @@
 #' @param aesthetics The ggplot2 aesthetics to which this scale should be applied.
 #' @param ... common discrete scale parameters: `name`, `breaks`, `labels`, `na.value`, `limits` and `guide`. See \code{\link[ggplot:discrete_scale]{discrete_scale}} for more details.
 #' @example ./examples/scale_color_c4a_cat.R
+#' @rdname scales_ggplot2
+#' @name scale_color_c4a_cat
 #' @export
-scale_color_c4a_cat = function (palette = NULL, reverse = NULL, order = NULL, aesthetics = "colour",
-		  ...)
-{
-	c4a_args <- c("palette", "reverse", "order")
-	args <- as.list(match.call())
-	args[[1]] <- NULL
-	args <- args[na.omit(match(c4a_args, names(args)))]
+scale_color_c4a_cat = function (palette = NULL, reverse = FALSE, order = NULL, ...) {
+	scale_discrete(aes = "colour", type = "cat", palette = palette, reverse = reverse, order = order, ...)
+}
+
+#' @rdname scales_ggplot2
+#' @name scale_colour_c4a_cat
+#' @export
+scale_colour_c4a_cat = function (palette = NULL, reverse = FALSE, order = NULL, ...) {
+	scale_discrete(aes = "colour", type = "cat", palette = palette, reverse = reverse, order = order, ...)
+}
+
+
+#' @rdname scales_ggplot2
+#' @name scale_fill_c4a_cat
+#' @export
+scale_fill_c4a_cat = function (palette = NULL, reverse = FALSE, order = NULL, ...) {
+	scale_discrete(aes = "fill", type = "cat", palette = palette, reverse = reverse, order = order, ...)
+}
+
+#' @rdname scales_ggplot2
+#' @name scale_color_c4a_seq
+#' @export
+scale_color_c4a_seq = function (palette = NULL, reverse = FALSE, contrast = NULL, ...) {
+	scale_discrete(aes = "color", type = "seq", palette = palette, reverse = reverse, contrast = contrast, ...)
+}
+
+#' @rdname scales_ggplot2
+#' @name scale_colour_c4a_seq
+#' @export
+scale_colour_c4a_seq = function (palette = NULL, reverse = FALSE, contrast = NULL, ...) {
+	scale_discrete(aes = "color", type = "seq", palette = palette, reverse = reverse, contrast = contrast, ...)
+}
+
[…]

### 2022-02-22 examples/scale_color_c4a_cat.R

@@ -1,12 +1,24 @@
 library("ggplot2")
 
 data("diamonds")
+diam_exp = diamonds[diamonds$price >= 15000, ]
+diam_exp$clarity[1:500] = NA
 
-# find suitable palette
-c4a_show(n = nlevels(diamonds$color), sort = "rank")
-
-# select "carto.safe"
-ggplot(diamonds[diamonds$price >= 15000, ], aes(x = carat, y = price, color = color)) +
+# discrete categorical scale
+ggplot(diam_exp, aes(x = carat, y = price, color = color)) +
 	geom_point(size = 2) +
 	scale_color_c4a_cat("carto.safe") +
+	theme_light()
+
+
+# missing values
+ggplot(diam_exp, aes(x = carat, y = price, fill = clarity)) +
+	geom_point(size = 2, shape = 21) +
+	scale_fill_c4a_cat("carto.safe") +
+	theme_light()
+
+# discrete sequential scale
+ggplot(diam_exp, aes(x = carat, y = price, color = cut)) +
+	geom_point(size = 2) +
+	scale_color_c4a_seq("scico.hawaii") +
 	theme_light()

### 2022-02-23 R/check_pals.R
improved ggplot2 scales: todo binned scales and docs

@@ -1,9 +1,9 @@
-#' Check diverging palette
-#'
-#' Check diverging palette. It computes two quality indices. \code{inter_wing_dist} minimal distance between from one wing (any color) to the other wing (any color); let a step be the distance from one color to a neighboring color, then: \code{min_step} is the minimal step. These two quality indices are computed for all three color vision deficiency types: per quality indicator, the worst score is returned.
-#'
-#' @param p diverging palette with odd number of colors (so with a middle color)
-#' @return vector of three quality indices
+# Check diverging palette
+#
+# Check diverging palette. It computes two quality indices. \code{inter_wing_dist} minimal distance between from one wing (any color) to the other wing (any color); let a step be the distance from one color to a neighboring color, then: \code{min_step} is the minimal step. These two quality indices are computed for all three color vision deficiency types: per quality indicator, the worst score is returned.
+#
+# @param p diverging palette with odd number of colors (so with a middle color)
+# @return vector of three quality indices
 check_div_pal = function(p) {
 	n = length(p)
 
@@ -54,14 +54,14 @@ check_div_pal = function(p) {
 }
 
 
-#' Check sequential palette
-#'
-#' Check sequential palette. It computes two quality indices. \code{min_step} and \code{max_step} are the minimum and maximum step respectively, where a step is the distance from one color to a neighboring color. \code{min_step} is the leading indicator: the higher, the better the palette. From palettes with equal \code{min_step}, those with the lowest \code{max_step} can be considered as better, because the steps are more uniform. These two quality indices are computed for all three color vision deficiency types: per quality indicator, the worst score is returned.
-#'
-#' and \code{max_step} is low, although the former is much more important.
-#'
-#' @param p sequential palette
-#' @return vector of three quality indices
+# Check sequential palette
+#
+# Check sequential palette. It computes two quality indices. \code{min_step} and \code{max_step} are the minimum and maximum step respectively, where a step is the distance from one color to a neighboring color. \code{min_step} is the leading indicator: the higher, the better the palette. From palettes with equal \code{min_step}, those with the lowest \code{max_step} can be considered as better, because the steps are more uniform. These two quality indices are computed for all three color vision deficiency types: per quality indicator, the worst score is returned.
+#
[…]

### 2022-02-23 R/ggplot2.R

@@ -2,90 +2,165 @@
 #'
 #' col4all scales for ggplot2
 #'
-#' @param palette,reverse,order See \code{\link{c4a}}.
-#' @param aesthetics The ggplot2 aesthetics to which this scale should be applied.
-#' @param ... common discrete scale parameters: `name`, `breaks`, `labels`, `na.value`, `limits` and `guide`. See \code{\link[ggplot:discrete_scale]{discrete_scale}} for more details.
-#' @example ./examples/scale_color_c4a_cat.R
+#' @param palette,reverse,order,contrast See \code{\link{c4a}}.
+#' @param mid data value that should be mapped to the mid-point of the diverging color scale
+#' @param n_interp number of discrete colors that should be used to interpolate the continuous color scale. Recommended to use an odd number to include the midpoint
+#' @param ... common discrete scale parameters: `name`, `breaks`, `labels`, `na.value`, `limits` and `guide`. See \code{\link[ggplot2:discrete_scale]{discrete_scale}} for more details.
+#' @example ./examples/scales_ggplot2.R
 #' @rdname scales_ggplot2
-#' @name scale_color_c4a_cat
+#' @name scale_color_discrete_c4a_cat
 #' @export
-scale_color_c4a_cat = function (palette = NULL, reverse = FALSE, order = NULL, ...) {
-	scale_discrete(aes = "colour", type = "cat", palette = palette, reverse = reverse, order = order, ...)
+scale_color_discrete_c4a_cat = function (palette = NULL, reverse = FALSE, order = NULL, ...) {
+	scale_discrete(aes = "color", type = "cat", palette = palette, reverse = reverse, order = order, ...)
 }
 
 #' @rdname scales_ggplot2
-#' @name scale_colour_c4a_cat
+#' @name scale_colour_discrete_c4a_cat
 #' @export
-scale_colour_c4a_cat = function (palette = NULL, reverse = FALSE, order = NULL, ...) {
-	scale_discrete(aes = "colour", type = "cat", palette = palette, reverse = reverse, order = order, ...)
+scale_colour_discrete_c4a_cat = function (palette = NULL, reverse = FALSE, order = NULL, ...) {
+	scale_discrete(aes = "color", type = "cat", palette = palette, reverse = reverse, order = order, ...)
 }
 
-
[…]

### 2022-02-23 examples/scale_color_c4a_cat.R

@@ -1,24 +1,0 @@
-library("ggplot2")
-
-data("diamonds")
-diam_exp = diamonds[diamonds$price >= 15000, ]
-diam_exp$clarity[1:500] = NA
-
-# discrete categorical scale
-ggplot(diam_exp, aes(x = carat, y = price, color = color)) +
-	geom_point(size = 2) +
-	scale_color_c4a_cat("carto.safe") +
-	theme_light()
-
-
-# missing values
-ggplot(diam_exp, aes(x = carat, y = price, fill = clarity)) +
-	geom_point(size = 2, shape = 21) +
-	scale_fill_c4a_cat("carto.safe") +
-	theme_light()
-
-# discrete sequential scale
-ggplot(diam_exp, aes(x = carat, y = price, color = cut)) +
-	geom_point(size = 2) +
-	scale_color_c4a_seq("scico.hawaii") +
-	theme_light()

### 2022-02-23 examples/scales_ggplot2.R

@@ -1,0 +1,36 @@
+library("ggplot2")
+
+data("diamonds")
+diam_exp = diamonds[diamonds$price >= 15000, ]
+diam_exp$clarity[1:500] = NA
+
+# discrete categorical scale
+ggplot(diam_exp, aes(x = carat, y = price, color = color)) +
+	geom_point(size = 2) +
+	scale_color_discrete_c4a_cat("carto.safe") +
+	theme_light()
+
+
+# missing values
+ggplot(diam_exp, aes(x = carat, y = price, fill = clarity)) +
+	geom_point(size = 2, shape = 21) +
+	scale_fill_discrete_c4a_cat("tol.muted") +
+	theme_light()
+
+# discrete sequential scale
+ggplot(diam_exp, aes(x = carat, y = price, color = cut)) +
+	geom_point(size = 2) +
+	scale_color_discrete_c4a_seq("hcl.blues2") +
+	theme_light()
+
+# continuous sequential scale
+ggplot(diam_exp, aes(x = carat, y = price, color = depth)) +
+	geom_point(size = 2) +
+	scale_color_continuous_c4a_seq("hcl.blues2", contrast = c(0.4, 1)) +
+	theme_light()
+
+# continuous sequential scale
+ggplot(diam_exp, aes(x = carat, y = depth, color = price)) +
+	geom_point(size = 2) +
+	scale_color_continuous_c4a_div("wes.zissou1", mid = mean(diam_exp$price)) +
+	theme_light()

### 2022-02-23 README.Rmd
improved ggplot2 scales: todo binned scales and docs
ready

@@
 "Color for all" refers to our mission that colors should be usable for not just people with normal color vision, but also for people with color vision deficiency.
-This package contains palettes from several popular color palette series, such as Color Brewer, Viridis, and Kovesi, but also the lesser known [palettes from Paul Tol](https://personal.sron.nl/~pault/).
+Currently, this package contains palettes from several popular and lesser known color palette series: [Color Brewer](https://colorbrewer2.org), [Viridis](http://bids.github.io/colormap/), [Kovesi](https://colorcet.com/), [Paul Tol](https://personal.sron.nl/~pault/), [Scico](https://www.fabiocrameri.ch/colourmaps/), [Carto](https://carto.com/carto-colors/), [Tableau](https://www.tableau.com/about/blog/2016/7/colors-upgrade-tableau-10-56782) and [Wes Anderson](https://github.com/karthik/wesanderson).
 Own palettes series can be added as well.
@@
 
-**cols4all** will be available on CRAN soon! Until then it can be installed using:
+**cols4all** will be available on CRAN soon. Until then it can be installed using:
 
@@
 library(cols4all)
+# For the GUI, shiny, shinyjs and kableExtra are required
+# install.packages(c("shiny", "shinyjs", "kableExtra"))
 c4a_gui()
@@
 
-Selecting a palette is easy. Palette names are in the format <series>.<palette_name>. So all letters are in small caps, and underscores are used to separate words.
+Selecting a palette is easy:
 
@@
 We use this package to analyse colors.
-For this purpose, we also use [**colorblindcheck**](https://github.com/Nowosad/colorblindcheck).
+For this purpose and specifically for color blind friendliness checks, we also use [**colorblindcheck**](https://github.com/Nowosad/colorblindcheck).
 
-There are a few features that distinguishes **cols4all** from other (great) color palette packages, in particular [**pals**](https://kwstat.github.io/pals/) and [**paletteer**](https://github.com/EmilHvitfeldt/paletteer):
+There are a few other pacakges with a large collection of color palettes, in particular [**pals**](https://kwstat.github.io/pals/) and [**paletteer**](https://github.com/EmilHvitfeldt/paletteer).
+There are a few features that distinguishes **cols4all** from those packages:
 
@@
 
-* With the gui (see below), a user is able to analyse characteristics of color palettes, such as color-blind-friendliness, harmony, and the presence of intense colors.
+* The GUI enables users to analyse characteristics of color palettes, such as color-blind-friendliness, harmony, and the presence of intense colors.
+
+* There is native support for **ggplot2** and **tmap** (as of the upcoming version 4).
 
@@
 
-There will be native support for **ggplot** and **tmap** (as of the upcoming version 4)
+## Feedback welcome!
+
+* Is everything working as expected?
+
+* Do you miss certain palettes or series?
+
+* Do you have ideas for improvement how to measure palette properties?
+
+Let us know! (via github issues)

### 2022-02-23 R/c4a_palettes.R

@@ -21,6 +21,7 @@ c4a_palettes = function(type = c("all", "cat", "seq", "div"), series = NULL, ful
 
 #' @rdname c4a_palettes
 #' @name c4a_series
+#' @export
 c4a_series = function(type = c("all", "cat", "seq", "div")) {
 	type = match.arg(type)
 	z = get(".z", envir = .C4A_CACHE)

### 2022-02-23 R/c4a_show.R

@@ -91,7 +91,7 @@ table_columns = function(type, show.scores) {
 #' @rdname c4a_gui
 #' @name c4a_gui
 c4a_show = function(type = c("cat", "seq", "div"), n = NULL, cvd.sim = c("none", "deutan", "protan", "tritan"), sort = "name", text.col = "same", series = NULL, contrast = NULL, include.na = FALSE, show.scores = FALSE, columns = NA) {
-	rm(id)
+	id = NULL
 	if (!requireNamespace("kableExtra")) stop("Please install kableExtra")
 
 	type = match.arg(type)

### 2022-02-23 README.Rmd
readme updated

@@
 
-The **cols4all** is an R package in development for selecting color palettes.
+The **cols4all** is a new R package for selecting color palettes.
 "Color for all" refers to our mission that colors should be usable for not just people with normal color vision, but also for people with color vision deficiency.
@@
 Own palettes series can be added as well.
-Color palettes are well organized.
+
+![categorical palettes)](https://user-images.githubusercontent.com/2444081/155185594-a52c361f-7113-40fe-82ea-94e782c27be4.png)
+![sequential palettes)](https://user-images.githubusercontent.com/2444081/155185615-9356443d-8a18-40a6-bc9e-d2e0d8e26eb4.png)
+![diverging palettes)](https://user-images.githubusercontent.com/2444081/155185638-3d05a045-c794-4eab-b994-88c0b58196e2.png)
+
+Color palettes are well organized and made consistent with each other.
+Moreover, they are scored on several aspects: color-blind-friendliness, the presence of intense colors (which should be avoided), the overall aesthetic harmony, and how many different hues are used.
+Finally, for each color palette a color for missing values is assigned, which is especially important for spatial data visualization.
 Currently we support three types: *categorical* (qualitative) palettes, *sequential* palettes, and *diverging* palettes.
-In the near future, more palette types will be added, namely *cyclic*, *bivariate*, and *hierarchical*.
+In the near future, more palette types will be added, such as *cyclic*, *bivariate*, and *hierarchical*.
 
@@
 
-Navigation through the color palettes is easy:
+Navigation through the color palettes is easy via the gui shown above, which is started with:
 
@@
 library(cols4all)
-# For the GUI, shiny, shinyjs and kableExtra are required
+# For the GUI, shiny, shinyjs, and kableExtra are required
 # install.packages(c("shiny", "shinyjs", "kableExtra"))
@@
 ```
-
-![categorical palettes)](https://user-images.githubusercontent.com/2444081/155185594-a52c361f-7113-40fe-82ea-94e782c27be4.png)
-![sequential palettes)](https://user-images.githubusercontent.com/2444081/155185615-9356443d-8a18-40a6-bc9e-d2e0d8e26eb4.png)
-![diverging palettes)](https://user-images.githubusercontent.com/2444081/155185638-3d05a045-c794-4eab-b994-88c0b58196e2.png)
 

### 2022-03-04 README.Rmd
c4a_ls added, copy icons improved

@@
 "Color for all" refers to our mission that colors should be usable for not just people with normal color vision, but also for people with color vision deficiency.
-Currently, this package contains palettes from several popular and lesser known color palette series: [Color Brewer](https://colorbrewer2.org), [Viridis](http://bids.github.io/colormap/), [Kovesi](https://colorcet.com/), [Paul Tol](https://personal.sron.nl/~pault/), [Scico](https://www.fabiocrameri.ch/colourmaps/), [Carto](https://carto.com/carto-colors/), [Tableau](https://www.tableau.com/about/blog/2016/7/colors-upgrade-tableau-10-56782) and [Wes Anderson](https://github.com/karthik/wesanderson).
+Currently, this package contains palettes from several popular and lesser known color palette series: [Color Brewer](https://colorbrewer2.org), [Viridis](http://bids.github.io/colormap/), [Kovesi](https://colorcet.com/), [Paul Tol](https://personal.sron.nl/~pault/), [Scico](https://www.fabiocrameri.ch/colourmaps/), [Carto](https://carto.com/carto-colors/), [Tableau](https://www.tableau.com/about/blog/2016/7/colors-upgrade-tableau-10-56782), [Wes Anderson](https://github.com/karthik/wesanderson), and [Seaborn](https://seaborn.pydata.org/tutorial/color_palettes.html).
 Own palettes series can be added as well.

### 2022-03-04 R/c4a_series_add.R

@@ -1,6 +1,6 @@
 #' Add and remove palette series
 #'
-#' Add and remove palette series. `c4a_series_add` can be used to add own palette series and  `c4a_series_remove` to remove palette series. `c4a_series_add_as_is` is the same as `c4a_series_add`, but by default without any processing.
+#' Add and remove palette series. `c4a_series_add` can be used to add own palette series and  `c4a_series_remove` to remove palette series. `c4a_series_add_as_is` is the same as `c4a_series_add`, but by default without any processing. These functions require the R package `colorblindcheck`.
 #'
 #' Indexing: for a categorical `"cat"` palette, an optional `"index"` attribute determines which colors to use for which lengths: if the palette consists of k colors, index should be a list of k, where the i-th element is an integer vector of length i with values 1,2,...,k. See `cols4all::.z$palette$rainbow` and  for an example.
 #'
@@ -113,6 +113,8 @@ c4a_series_add = function(x, xNA = NA, types, series, format.palette.name = TRUE
 
 	.C4A$z = z
 	.C4A$s = s
+	fill_ls()
+	invisible(NULL)
 }
 
 #' @rdname c4a_series_add
@@ -183,4 +185,6 @@ c4a_series_remove = function(series = "all", are.you.sure = FALSE) {
 	}
 	.C4A$z = z2
 	.C4A$s = s2
+	fill_ls()
+	invisible(NULL)
 }

### 2022-03-04 R/c4a_sysdata_import.R

@@ -19,6 +19,8 @@ c4a_sysdata_import = function(x) {
 
 	.C4A$z = z
 	.C4A$s = s
+	fill_ls()
+	invisible(NULL)
 }
 
 #' @rdname c4a_sysdata_import

### 2022-03-04 R/c4a_table.R

@@ -30,7 +30,7 @@ table_columns = function(type, show.scores) {
 
 #' Graphical user interface to analyse palettes
 #'
-#' Graphical user interface to analyse palettes. `c4a_table` shows a table that can be opened in the browser. `c4a_gui` is a graphical user interface (shiny app) around this table.
+#' Graphical user interface to analyse palettes. `c4a_table` shows a table that can be opened in the browser. `c4a_gui` is a graphical user interface (shiny app) around this table. The package `kableExtra` is required for `c4a_table` and for `c4a_gui` the pacakges `shiny` and `shinyjs`.
 #'
 #' @section Table Columns:
 #'
@@ -84,7 +84,6 @@ table_columns = function(type, show.scores) {
 #' @param sort column name to sort the data. For column names, see details. Use a `"-"` prefix to reverse the order.
 #' @param text.format The format of the text of the colors. One of `"hex"`, `"RGB"` or `"HCL"`.
 #' @param text.col The text color of the colors. By default `"same"`, which means that they are the same as the colors themselves (so invisible, but available for selection).
-#' @param text.copy Copy format. `"R"` stands for `c("#111111", "#222222")` and `"other"` stands for `["#111111", "#222222"]`.
 #' @param series Series of palettes to show. See \code{\link{c4a_series}} for options. By default, `"all"`, which means all series. For `c4a_gui` it only determines which series are shown initially.
[…]

### 2022-03-04 R/onLoad.R

@@ -48,6 +48,27 @@
 
 		nmax = c(cat = 36, seq = 15, div = 15)
 	})
+	fill_ls()
 }
 
 .C4A <- new.env(FALSE, parent=globalenv())
+
+
+#' Nested list of cols4all palette names
+#'
+#' Nested list of cols4all palette names, which is handy for auto-completion
+#'
+#' @export
+c4a_ls <- new.env(FALSE, parent=globalenv())
+
+
+fill_ls = function() {
+	z = .C4A$z
+	x = sort(unique(z$series))
+	y = structure(lapply(x, function(xi) {
+		zi = z[z$series == xi, ]
+		structure(as.list(zi$fullname), names = zi$name)
+	}), names = x)
+	list2env(y, envir = c4a_ls)
+}
+

### 2022-03-04 README.Rmd
updated docs, added c4a_ls

@@
 
-# Colors for all!
+# Colors for all! <img src='inst/img/cols4all_logo.png' align="right" height="139" />
 
@@
 
-![categorical palettes)](https://user-images.githubusercontent.com/2444081/155185594-a52c361f-7113-40fe-82ea-94e782c27be4.png)
-![sequential palettes)](https://user-images.githubusercontent.com/2444081/155185615-9356443d-8a18-40a6-bc9e-d2e0d8e26eb4.png)
-![diverging palettes)](https://user-images.githubusercontent.com/2444081/155185638-3d05a045-c794-4eab-b994-88c0b58196e2.png)
+Categorical             |  Sequential | Diverging
+:-------------------------:|:-------------------------:|:-------------------------:
+[![](https://user-images.githubusercontent.com/2444081/156847689-7e1a4608-9b03-45f5-9ccf-56e6d8844d47.png)](https://user-images.githubusercontent.com/2444081/156847689-7e1a4608-9b03-45f5-9ccf-56e6d8844d47.png) | [![](https://user-images.githubusercontent.com/2444081/156847703-b429c521-89d5-4e14-85c1-0f100ca93b45.png)](https://user-images.githubusercontent.com/2444081/156847703-b429c521-89d5-4e14-85c1-0f100ca93b45.png) | [![](https://user-images.githubusercontent.com/2444081/156847712-33a3cd22-a50b-44e2-9666-b67fcb0d2853.png)]((https://user-images.githubusercontent.com/2444081/156847712-33a3cd22-a50b-44e2-9666-b67fcb0d2853.png))
 
@@
 
-```{r}
+```{r, echo=FALSE}
 getwd()
@@
 
-## Get started
+## Getting started
 
-Navigation through the color palettes is easy via the gui shown above, which is started with:
+Navigation through the color palettes is easy via the GUI shown above, which is started with:
 
@@
 ```
+
+Besides browsing through palettes, it is also easy to copy color codes to the clipboard, either by selecting the hidden text in the color tables or by clicking on the icons at the right-hand-side:
+
+![](https://user-images.githubusercontent.com/2444081/156849917-59af1dc1-ee5e-40be-bf87-5c7384351ae0.png){width=50%}
 
@@
 ```
+
+## Overview of functions
+
+Main functions:
+
+* `c4a_gui` GUI (shiny app) to see and analyse the palettes
+* `c4a` Get the colors of a palette
+
+Palette names and properties:
+
+* `c4a_palettes` Get available palette names
+* `c4a_series` Get available series
+* `c4a_meta` Get meta information (such as type and maximum number of colors )
+* `c4a_ls` Environment via which palette names can be browsed with auto-completion (using `$`)
+
+Importing and exporting palettes:
+
+* `c4a_series_add` Add color palettes
+* `c4a_series_remove` Remove color palettes
+* `c4a_sysdata_import` Import system data
+* `c4a_sysdata_export` Export system data
+
+ggplot2
+
+* `scale_<aesthetic>_<mapping>_c4a_<type>` e.g. `scale_color_continuous_c4a_div` Add scale to ggplot2.
 

### 2022-03-04 R/c4a.R

@@ -8,6 +8,7 @@
 #' @param reverse should the palette be reversed?
 #' @param order order of colors. Only applicable for `"cat"` palettes
 #' @param contrast a vector of two numbers between 0 and 1 that determine the range that is used for sequential and diverging palettes. The first number determines where the palette begins, and the second number where it ends. For sequential `"seq"` palettes, 0 means the leftmost (normally lightest) color, and 1 the rightmost (often darkest) color. For diverging `"seq"` palettes, 0 means the middle color, and 1 both extremes. If only one number is provided, this number is interpreted as the endpoint (with 0 taken as the start). The default values (that depend on the `n`n and `type`) are provided by \code{\link{c4a_default_contrast}}.
+#' @param format format of the colors. One of: `"hex"` character vector of hex color values, `"RGB"` 3 column matrix of RGB values, or `"HCL"` 3-column matrix of HCL values
 #' @param n_too_large what should be done in case `n` is larger than the maximum amount of colors? Options are `"error"` (an error is returned), `"repeat"`, the palette is repeated, `"interpolate"` colors are interpolated. For categorical `"cat"` palettes only.
 #' @param verbose should messages be printed?
 #' @return A vector of colors
@@ -18,8 +19,10 @@
 #' @rdname c4a
 #' @name c4a
 #' @export
-c4a = function(palette = NULL, n = NULL, type = c("cat", "seq", "div"), reverse = FALSE, order = NULL, contrast = NA, n_too_large = c("error", "repeat", "interpolate"), verbose = TRUE) {
+c4a = function(palette = NULL, n = NULL, type = c("cat", "seq", "div"), reverse = FALSE, order = NULL, contrast = NA, format = c("hex", "RGB", "HCL"), n_too_large = c("error", "repeat", "interpolate"), verbose = TRUE) {
 	type = match.arg(type)
+	format = match.arg(format)
+
 	n_too_large = match.arg(n_too_large)
 
 	if (is.null(palette)) {
@@ -47,7 +50,15 @@ c4a = function(palette = NULL, n = NULL, type = c("cat", "seq", "div"), reverse 
 	} else pal
 
 	if (!is.null(mes) && verbose) message(mes)
-	if (reverse) rev(pal) else pal
+	pal2 = if (reverse) rev(pal) else pal
+
+	if (format == "hex") {
+		pal2
+	} else if (format == "RGB") {
[…]

### 2022-03-04 R/c4a_gui.R

@@ -20,6 +20,8 @@ c4a_gui = function(type = "cat", n = 9, series = "all") {
 		return(invisible(NULL))
 	}
 
+	shiny::addResourcePath(prefix = "imgResources", directoryPath = "inst/img")
+
 	ui = shiny::fluidPage(
 		#shinyjs::useShinyjs(),
 		shiny::tags$style(shiny::HTML("div.sticky {
@@ -30,12 +32,20 @@ c4a_gui = function(type = "cat", n = 9, series = "all") {
 		}")),
 
 		# Application title
-		shiny::titlePanel("col4all: colors for all!"),
+		#shiny::titlePanel("col4all: colors for all!"),
+
+		shiny::titlePanel(title = "Colors for all!"),
+
 
 		shiny::sidebarLayout(
 			shiny::tagAppendAttributes(shiny::sidebarPanel(
 				width = 3,
-				shiny::radioButtons("type", "Type", choices = c(Categorical = "cat", Sequential = "seq", Diverging = "div"), selected = type), #, Cyclic = "cyc", Bivariate = "biv", Tree = "tree"), selected = "cat"),
+				shiny::fluidRow(
+				shiny::column(9,
+					shiny::radioButtons("type", "Type", choices = c(Categorical = "cat", Sequential = "seq", Diverging = "div"), selected = type)
+				), shiny::column(3,
+					shiny::img(align = "right", alt = "test", src = "imgResources/cols4all_logo.png", height = 100)
+				)),
 				shiny::selectizeInput("series", "Palette Series", choices = allseries, selected = series, multiple = TRUE),
 				shiny::sliderInput("n", "Number of colors",
 								   min = 2, max = 36, value = n, ticks = FALSE),

### 2022-03-04 R/c4a_palettes.R

@@ -1,6 +1,6 @@
-#' List all available cols4all color palettes and series
+#' Get available palette names and series
 #'
-#' `c4a_palettes` lists all available cols4all color palettes. Palettes are organized by series. The available series are listed with `c4a_series`. The references are provided below. Palettes are also organized per functional type, where we currently support: categorical `"cat"`, sequential `"seq"`, and diverging `"div"`" palette types. The function `c4a_default_palette` returns the default (recommended) palette per type.
+#' `c4a_palettes` lists all available cols4all color palettes. Palettes are organized by series. The available series are listed with `c4a_series`. In an IDE with auto-completion (such as RStudio) it is possible to browse through the palette names with `c4a_ls` (using `$` like in lists). Palettes are also organized per functional type, where we currently support: categorical `"cat"`, sequential `"seq"`, and diverging `"div"`" palette types. The function `c4a_default_palette` returns the default (recommended) palette per type. The references are provided below.
 #'
 #' @section References:
 #'
@@ -79,6 +79,10 @@
 #'
 #' Obtained via the R package `ggthemes`
 #'
+#' \strong{seaborn}
+#'
+#' Palettes by Python library `seaborn` by Michael Waskom. \url{https://seaborn.pydata.org/tutorial/color_palettes.html}
+#'
 #' @param type type of color palette: one of `"all"` (all palettes), `"cat"` (categorical/qualitative palettes), `"seq"` (sequential palettes) and `"div"` (diverging palettes).
 #' @param series series to list the palettes from. Run `c4a_series` to see the options.
 #' @param full.names should full names, i.e. with the prefix "series."? By default `TRUE`.
@@ -113,3 +117,9 @@ c4a_default_palette = function(type = c("cat", "seq", "div")) {
 	type = match.arg(type)
 	.C4A$defaults[type]
 }
+
+
+#' @rdname c4a_palettes
+#' @name c4a_ls
+#' @export
+c4a_ls <- new.env(FALSE, parent=globalenv())

### 2022-03-04 R/check_pals.R

@@ -29,10 +29,10 @@ check_div_pal = function(p) {
 					tri = colorspace::tritan(p),
 					p)
 		m = get_hcl_matrix(p2)
-		m[,3][m[,2]<10] = NA
+		m[,1][m[,2]<10] = NA
 
-		h1 = m[1:nh, 3]
-		h2 = m[(nh+1+!is_even):n, 3]
+		h1 = m[1:nh, 1]
+		h2 = m[(nh+1+!is_even):n, 1]
 
 		inter_wing_hue_dist = if (all(is.na(h1)) && all(is.na(h2))) {
 			0
@@ -117,12 +117,12 @@ check_cat_pal = function(p) {
 
 # get hcl coordinates
 get_hcl_matrix = function(p, rounded = FALSE) {
-	x = as(hex2RGB(p), "polarLUV")@coords
+	x = as(hex2RGB(p), "polarLUV")@coords[,c("H", "C", "L")]
 	if (rounded) round(x) else x
 }
 
 get_hcl_triple = function(p) {
-	x = get_hcl_matrix(p, rounded = TRUE)[,c("H", "C", "L")]
+	x = get_hcl_matrix(p, rounded = TRUE)
 	apply(x, MARGIN = 1, paste, collapse = ",")
 }
 
@@ -163,7 +163,7 @@ analyse_hcl = function(p) {
 	m = get_hcl_matrix(p)
 
 	# hue width: how far are hues apart from each other?
-	h = round(m[,3])
+	h = round(m[,1])
 	h[m[,2]<=.C4A$Cgray] = NA
 
 	Hwidth = get_hue_width(h)
@@ -181,7 +181,7 @@ analyse_hcl = function(p) {
 	HwidthR = get_hue_width(hR)
 
 	Cmax = round(max(m[,2]))
-	Lrange = round(max(m[,1]) - min(m[,1]))
+	Lrange = round(max(m[,3]) - min(m[,3]))
 	Crange = round(max(m[,2]) - min(m[,2]))
 	LCrange = round(max(Lrange * .C4A$LrangeWeight, Crange * (1-.C4A$LrangeWeight)))
 

### 2022-03-04 R/onLoad.R

@@ -1,3 +1,40 @@
+#' cols4all
+#'
+#' Color palettes for all people, including those with color vision deficiency. Palettes, organized by source and type, and scores on several properties such as color-blind-friendliness and harmony. Furthermore each palette is assigned a distinct color for missing values. Own palettes can be loaded as well. ggplot2 scales are included.
+#'
+#' This page provides a brief overview of all package functions.
+#'
+#' @section Main functions:
+#' \tabular{ll}{
+#' \code{\link{c4a_gui}}\tab GUI (shiny app) to see and analyse the palettes \cr
+#' \code{\link{c4a}}\tab Get the colors of a palette \cr
+#' }
+#'
+#' @section Palette names and properties:
+#' \tabular{ll}{
+#' \code{\link{c4a_palettes}}\tab Get available palette names \cr
+#' \code{\link{c4a_series}}\tab Get available series \cr
+#' \code{\link{c4a_meta}}\tab Get meta information (such as type and maximum number of colors ) \cr
+#' \code{\link{c4a_ls}}\tab Environment via which palette names can be browsed with auto-completion (using `$`) \cr
+#' }
+#'
+#' @section Importing and exporting palettes:
+#' \tabular{ll}{
+#' \code{\link{c4a_series_add}}\tab Add color palettes \cr
+#' \code{\link{c4a_series_remove}}\tab Remove color palettes \cr
+#' \code{\link{c4a_sysdata_import}}\tab Import system data \cr
+#' \code{\link{c4a_sysdata_export}}\tab Export system data \cr
+#' }
+#'
+#' @name cols4all-package
+#' @aliases cols4all
+#' @docType package
+#' @author Martijn Tennekes \email{mtennekes@@gmail.com}
+#' @seealso \url{https://github.com/mtennekes/cols4all}
+#' @concept color
+#' @concept visualization
+NULL
+
 .onLoad <- function(...) {
 	assign("z", .z, envir = .C4A)
 	assign("s", .s, envir = .C4A)
@@ -53,13 +90,6 @@
 
 .C4A <- new.env(FALSE, parent=globalenv())
 
-
-#' Nested list of cols4all palette names
-#'
-#' Nested list of cols4all palette names, which is handy for auto-completion
-#'
-#' @export
-c4a_ls <- new.env(FALSE, parent=globalenv())
 
 
 fill_ls = function() {

### 2022-03-04 R/process_palette.R

@@ -11,7 +11,7 @@ process_palette = function(pal, type, colNA = NA, take.gray.for.NA = TRUE, remov
 	#specplot(hcl(h=seq(0,360,by=10), c = 10, l= 5))
 	#specplot(hcl(h=seq(0,360,by=10), c = 15, l= 0))
 	if (remove.blacks && type == "cat") {
-		isB = (hcl[,1] + hcl[,2]) <= 15
+		isB = (hcl[,3] + hcl[,2]) <= 15
 		if (any(isB)) {
 			pal = pal[!isB]
 			hcl = hcl[!isB,]
@@ -23,7 +23,7 @@ process_palette = function(pal, type, colNA = NA, take.gray.for.NA = TRUE, remov
 		if (take.gray.for.NA) {
 			wG = which(hcl[,2] < .C4A$Cgray)
 			if (length(wG) && is.na(colNA)) {
-				wNA = wG[which.max(hcl[wG, 1])]
+				wNA = wG[which.max(hcl[wG, 3])]
 				colNA = pal[wNA]
 				pal = pal[-wNA]
 				hcl = hcl[-wNA,]
@@ -42,7 +42,7 @@ process_palette = function(pal, type, colNA = NA, take.gray.for.NA = TRUE, remov
 	}
 
 	if (light.to.dark && type == "seq") {
-		ls = hcl[,1]
+		ls = hcl[,3]
 		ls_sg = sign(ls[-1] - ls[-length(ls)])
 		if (all(ls_sg>=0)) {
 		#if ((hcl[nrow(hcl), 1] - hcl[1,1]) > 30) {
@@ -62,7 +62,7 @@ process_palette = function(pal, type, colNA = NA, take.gray.for.NA = TRUE, remov
 		# first candidates: choose NA from grays, such that luminance is at most 0.3 lighter and not darker than the lightest resp. darkest color.
 		# prefer lightest gray
 
-		gray_range = c(min(hcl[,1]/100), min(1, (max(hcl[,1]/100) + 0.3)))
+		gray_range = c(min(hcl[,3]/100), min(1, (max(hcl[,3]/100) + 0.3)))
 		candidates = list(grDevices::gray.colors(10, start = gray_range[1], end = gray_range[2]),
 						  grDevices::hcl(h = seq(0, 340, by = 20), c = 30, l = 70),
 						  grDevices::hcl(h = seq(0, 340, by = 20), c = 50, l = 70))

### 2022-03-29 README.Rmd
added range_matrix and space attributes, removed default range, rebuild palettes

@@
 Finally, for each color palette a color for missing values is assigned, which is especially important for spatial data visualization.
-Currently we support three types: *categorical* (qualitative) palettes, *sequential* palettes, and *diverging* palettes.
-In the near future, more palette types will be added, such as *cyclic*, *bivariate*, and *hierarchical*.
+Currently we support several types: *categorical* (qualitative) palettes, *sequential* palettes, *diverging* palettes, and *bivariate* palettes (divided into three types).
+In the near future, more palette types may be added, such as *cyclic* and *hierarchical*.
 
@@
 
-* `c4a_series_add` Add color palettes
-* `c4a_series_remove` Remove color palettes
+* `c4a_palettes_add` Add color palettes
+* `c4a_palettes_remove` Remove color palettes
 * `c4a_sysdata_import` Import system data

### 2022-03-29 R/c4a.R

@@ -8,7 +8,7 @@
 #' @param type type of color palette, in case `palette` is not specified: one of `"cat"` (categorical/qualitative palette), `"seq"` (sequential palette), `"div"` (diverging palette), and `"bivs"`/`"bivc"`/`"bivu"` (bivariate: respectively seq-seq cat-seq, and uncertainty-seq).
 #' @param reverse should the palette be reversed?
 #' @param order order of colors. Only applicable for `"cat"` palettes
-#' @param range a vector of two numbers between 0 and 1 that determine the range that is used for sequential and diverging palettes. The first number determines where the palette begins, and the second number where it ends. For sequential `"seq"` palettes, 0 means the leftmost (normally lightest) color, and 1 the rightmost (often darkest) color. For diverging `"seq"` palettes, 0 means the middle color, and 1 both extremes. If only one number is provided, this number is interpreted as the endpoint (with 0 taken as the start). The default values (that depend on the `n`n and `type`) are provided by \code{\link{c4a_default_range}}.
+#' @param range a vector of two numbers between 0 and 1 that determine the range that is used for sequential and diverging palettes. The first number determines where the palette begins, and the second number where it ends. For sequential `"seq"` palettes, 0 means the leftmost (normally lightest) color, and 1 the rightmost (often darkest) color. For diverging `"seq"` palettes, 0 means the middle color, and 1 both extremes. If only one number is provided, this number is interpreted as the endpoint (with 0 taken as the start).
 #' @param format format of the colors. One of: `"hex"` character vector of hex color values, `"RGB"` 3 column matrix of RGB values, or `"HCL"` 3-column matrix of HCL values
 #' @param n_invalid what should be done in case `n` is larger than the maximum number of colors or smaller than the minimum number? Options are `"error"` (an error is returned), `"repeat"`, the palette is repeated, `"interpolate"` colors are interpolated. For categorical `"cat"` palettes only.
 #' @param verbose should messages be printed?

### 2022-03-29 R/c4a_default_contrast.R

@@ -1,48 +1,56 @@
-#' Get the default range
-#'
-#' Get the default range for gradient palettes, so sequential `"seq"` and diverging `"div"`. The range is a vector of two numbers between 0 and 1 that determine the range that is used for sequential and diverging palettes. The first number determines where the palette begins, and the second number where it ends. For sequential `"seq"` palettes, 0 means the leftmost (normally lightest) color, and 1 the rightmost (often darkest) color. For diverging `"seq"` palettes, 0 means the middle color, and 1 both extremes. If only one number is provided, this number is interpreted as the endpoint (with 0 taken as the start). This function sets the default based on the number of colors `n`: for small values of `n` the default range is smaller than `c(0, 1)` for aesthetically more pleasing palettes. See `c4a_gui` how the range effects the palettes. See also examples below. The range can be set manually with the argument `range` of \code{\link{c4a}}.
-#'
-#' @param n number of colors
-#' @param type type of color palette: one of `"seq"` (sequential palettes), `"div"` (diverging palettes), and `"cat"` (categorical/qualitative palettes).
-#' @example ./examples/c4a_default_range.R
-#' @export
-c4a_default_range = function(n, type = c("seq", "div", "bivs", "bivc", "bivu", "cat")) {
-	type = match.arg(type)
-	fun = paste0("default_range_", type)
-	do.call(fun, list(k = n))
-}
+# Get the default range
+#
[…]

### 2022-03-29 R/c4a_gui.R

@@ -22,7 +22,7 @@ c4a_gui = function(type = "cat", n = NA, series = c("misc", "brewer", "hcl", "to
 		series = intersect(series, allseries)
 	}
 	if (!length(series)) {
-		message("No palette series loaded. Please reload cols4all, add series with c4a_series_add, or import data with c4a_sysdata_import")
+		message("No palette series loaded. Please reload cols4all, add series with c4a_palettes_add, or import data with c4a_sysdata_import")
 		return(invisible(NULL))
 	}
 
@@ -105,11 +105,13 @@ c4a_gui = function(type = "cat", n = NA, series = c("misc", "brewer", "hcl", "to
 					shiny::fluidRow(
 						shiny::column(4,
 							#shiny::br(),
-							shiny::radioButtons("auto_range", label = "Range", choices = c("Maximum", "Automatic", "Manual"), selected = "Maximum")),
+							shiny::radioButtons("auto_range", label = "Range", choices = c("Automatic", "Manual"), selected = "Automatic")),
+						shiny::conditionalPanel(
+							condition = "input.auto_range == 'Manual'",
 							shiny::column(8,
-								shinyjs::disabled(shiny::div(style = "font-size:0;margin-bottom:-20px", shiny::sliderInput("range", "",
-												   min = 0, max = 1, value = c(0,1), step = .05))),
-								shiny::uiOutput("range_info")))),
+								shiny::div(style = "font-size:0;margin-bottom:-20px", shiny::sliderInput("range", "",
+												   min = 0, max = 1, value = c(0,1), step = .05)),
+								shiny::uiOutput("range_info"))))),
 				shiny::fluidRow(
 					shiny::column(7,
 								  shiny::radioButtons("format", "Text format", choices = c("Hex" = "hex", "RGB" = "RGB", "HCL" = "HCL"), inline = TRUE)),
@@ -123,9 +125,9 @@ c4a_gui = function(type = "cat", n = NA, series = c("misc", "brewer", "hcl", "to
 					shiny::column(5,
 						shiny::br(),
 						shiny::checkboxInput("sortRev", "Reverse", value = FALSE))), style = "margin-bottom:-10px;"),
-				shiny::checkboxInput("dark", "Dark mode", value = FALSE),
-				shiny::checkboxInput("advanced", "Show underlying scores", value = FALSE)
[…]

### 2022-03-29 R/c4a_palettes_add.R

@@ -1,61 +1,83 @@
 #' Add and remove palette series
 #'
-#' Add and remove palette series. `c4a_series_add` can be used to add own palette series and  `c4a_series_remove` to remove palette series. `c4a_series_add_as_is` is the same as `c4a_series_add`, but by default without any processing. These functions require the R package `colorblindcheck`.
+#' Add and remove palette series. `c4a_palettes_add` can be used to add own palette series and  `c4a_palettes_remove` to remove palette series. `c4a_palettes_add_as_is` is the same as `c4a_palettes_add`, but by default without any processing. These functions require the R package `colorblindcheck`.
 #'
-#' Indexing: for a categorical `"cat"` palette, an optional `"index"` attribute determines which colors to use for which lengths: if the palette consists of k colors, index should be a list of k, where the i-th element is an integer vector of length i with values 1,2,...,k. See `cols4all::.z$palette$rainbow` and  for an example.
+#' Indexing: for a categorical `"cat"` palette, an optional `"index"` attribute determines which colors to use for which lengths: if the palette consists of k colors, index should be a list of k, where the i-th element is an integer vector of length i with values 1,2,...,k. See `c4a_meta("rainbow")` and  for an example.
 #'
-#' It may take some time to process, especially large categorical palettes, because of calculations of the color blind checks
+#' Range: sequential and diverging palettes are usually defined for 9+ colors. The optional `"range_matrix"` attribute determines that range is used for less colors. It is a n x 2 matrix where row i defines the applied range of a palette of length i. For sequential palettes a range `c(0,1)` means that the palette is generated (via a color ramp) between the two outermost colors. For diverging palettes, a range `c(x, y)` means that both sides of the palette are generated (via a color ramp) from `x`, which is the distance to the center color, to `y` which represents both outermost colors.
+#'
[…]

### 2022-03-29 R/c4a_sysdata_import.R

@@ -1,6 +1,6 @@
 #' Import and export system data
 #'
-#' Import and export system data. `c4a_sysdata_import` will import system data and overwrite the current system data. System data includes palette data and pre-calculated color-blind-friendliness scores. `c4a_sysdata_export` will export the current system data. These functions are for advanced use; the functions \code{\link{c4a_series_add}} and \code{\link{c4a_series_remove}} are user-friendly interfaces to change the loaded palettes.
+#' Import and export system data. `c4a_sysdata_import` will import system data and overwrite the current system data. System data includes palette data and pre-calculated color-blind-friendliness scores. `c4a_sysdata_export` will export the current system data. These functions are for advanced use; the functions \code{\link{c4a_palettes_add}} and \code{\link{c4a_palettes_remove}} are user-friendly interfaces to change the loaded palettes.
 #'
 #' @param x system data
 #' @example ./examples/c4a_sysdata_import.R

### 2022-03-29 R/c4a_table.R

@@ -125,18 +125,9 @@ c4a_table = function(type = c("cat", "seq", "div", "bivs", "bivc", "bivu"), n = 
 	z = .C4A$z
 
 	if (is.null(z)) {
-		if (verbose) message("No palette series loaded. Please reload cols4all, add series with c4a_series_add, or import data with c4a_sysdata_import")
+		if (verbose) message("No palette series loaded. Please reload cols4all, add series with c4a_palettes_add, or import data with c4a_sysdata_import")
 		return(invisible(NULL))
 	}
-
-	if (is.na(range[1])) {
-		if (!is.null(n)) {
-			range = c4a_default_range(n, type)
-		} else {
-			range = c(0, 1)
-		}
-	}
-
 
 	zn = get_z_n(z[z$type == type, ], n = n, m = m, range = range)
 	if (!is.null(zn)) {
@@ -148,7 +139,7 @@ c4a_table = function(type = c("cat", "seq", "div", "bivs", "bivc", "bivu"), n = 
 		return(invisible(NULL))
 	}
 
-	zn = show_attach_scores(zn, range = range)
+	zn = show_attach_scores(zn)
 	# better but slower alternative: calculate all scores read time:
 	#zn = get_scores_zn(zn)
 

### 2022-03-29 R/get_pal_n.R

@@ -15,10 +15,8 @@ get_pal_n = function(n, m = NA, name, type, series, palette, nmin, nmax, ndef, r
 		n = ndef
 	}
 	index = attr(palette, "index")
-
-	if (is.na(range[1])) {
-		range = c4a_default_range(n, type)
-	}
+	range_matrix = attr(palette, "range_matrix")
+	space = attr(palette, "space")
 
 	x = if (type == "cat") {
 		if (is.null(index)) {
@@ -26,33 +24,56 @@ get_pal_n = function(n, m = NA, name, type, series, palette, nmin, nmax, ndef, r
 		} else {
 			palette[index[[n]]]
 		}
-	} else if (type == "seq") {
-		if (range[1] == 0 && range[2] == 1) {
-			rampPal(palette, n)
+	} else if (type %in% c("seq", "div")) {
+		if (is.na(range[1])) {
+			if (!is.null(index)) {
+				pal = palette[index[[min(n, length(index))]]]
+				rng = c(0, 1)
+			} else if (!is.null(range_matrix)) {
+				pal = palette
+				rng = range_matrix[min(n, nrow(range_matrix)), ]
+			} else {
+				pal = palette
+				rng = c(0, 1)
+			}
 		} else {
-			rangeIDs <- round(seq(range[1]*100, range[2]*100, length.out=n))+1
-			rampPal(palette, 101)[rangeIDs]
+			if (!is.null(index)) {
+				pal = palette[index[[length(index)]]]
+			} else {
+				pal = palette
+			}
+			rng = range
 		}
-	} else if (type == "div") {
-		if (range[1] == 0 && range[2] == 1) {
-			rampPal(palette, n)
+
+		if (type == "seq") {
+			if (rng[1] == 0 && rng[2] == 1) {
+				rampPal(pal, n, space = space)
+			} else {
+				rngIDs <- round(seq(rng[1]*100, rng[2]*100, length.out=n))+1
+				rampPal(pal, 101, space = space)[rngIDs]
+			}
 		} else {
-			breaks = seq(-10,10, length.out=n+1)
-			rangeIDs <- map2divscaleID(breaks=breaks, range=range)
-			rampPal(palette, 101)[rangeIDs]
+			if (rng[1] == 0 && rng[2] == 1) {
+				rampPal(pal, n, space = space)
+			} else {
+				breaks = seq(-10,10, length.out=n+1)
[…]

### 2022-03-29 R/onLoad.R

@@ -20,8 +20,8 @@
 #'
 #' @section Importing and exporting palettes:
 #' \tabular{ll}{
-#' \code{\link{c4a_series_add}}\tab Add color palettes \cr
-#' \code{\link{c4a_series_remove}}\tab Remove color palettes \cr
+#' \code{\link{c4a_palettes_add}}\tab Add color palettes \cr
+#' \code{\link{c4a_palettes_remove}}\tab Remove color palettes \cr
 #' \code{\link{c4a_sysdata_import}}\tab Import system data \cr
 #' \code{\link{c4a_sysdata_export}}\tab Export system data \cr
 #' }

### 2022-03-29 R/process_palette.R

@@ -1,7 +1,8 @@
-process_palette = function(pal, type, colNA = NA, take.gray.for.NA = TRUE, remove.other.grays = FALSE, remove.blacks = TRUE, light.to.dark = TRUE, remove.names = TRUE, biv.method = "byrow") {
+process_palette = function(pal, type, colNA = NA, take.gray.for.NA = TRUE, remove.other.grays = FALSE, remove.blacks = TRUE, light.to.dark = TRUE, remove.names = TRUE, biv.method = "byrow", space = "rgb", range_matrix_args = list()) {
 
 	# maybe need to reindex
 	index = attr(pal, "index")
+	range_matrix = attr(pal, "range_matrix")
 	orig_pal = pal
 
 	if (type %in% c("bivs", "bivc", "bivu")) {
@@ -110,7 +111,24 @@ process_palette = function(pal, type, colNA = NA, take.gray.for.NA = TRUE, remov
 			index2[[w]]
 		})
 		attr(pal, "index") = index3
+	} else if (is.null(range_matrix) && type %in% c("seq", "div")) {
+
+		rma = formals(get(paste0("range_", type)))
+		rma$n = NULL
+
+
+		if (!is.null(range_matrix_args) && length(range_matrix_args)) {
+			rma[names(range_matrix_args)] = range_matrix_args
+		}
+
+		range_matrix = do.call(paste0("range_", type), c(list(n = length(pal)), rma))
+		attr(pal, "range_matrix") = range_matrix
+	} else if (is.null(range_matrix)) {
+		attr(pal, "range_matrix") = range_matrix
 	}
+
+	attr(pal, "space") = space
+
 	pal[] = toupper(pal[])
 	colNA = toupper(colNA)
 

### 2022-03-29 R/validate_colors.R

@@ -21,9 +21,11 @@ validate_colors <- function(x) {
 		index = attr(x, "index")
 		if (!is.list(index)) stop("index attribute is not a list")
 		if (length(index) > length(x)) stop("index too large")
+		if (is.null(names(index))) names(index) = 1:length(index)
 		for (i in 1:length(index)) {
 			ind = index[[i]]
-			if (!is.numeric(ind) || any(is.na(ind)) || any(ind > length(x)) || length(ind) != i) stop("Incorrect index numbers", call. = FALSE)
+			len = as.character(names(index)[i])
+			if (!is.null(ind) && (!is.numeric(ind) || any(is.na(ind)) || any(ind > length(x)) || length(ind) != len)) stop("Incorrect index numbers", call. = FALSE)
 		}
 	}
 	x

### 2022-03-29 build/build_palettes.R

@@ -12,12 +12,24 @@ library(ggthemes)
 library(reticulate) # to get seaborn
 library(Polychrome)
 
-## more:
+devtools::session_info(pkgs = "attached")
+# ! package         * version date (UTC) lib source
+# colorspace      * 2.0-3   2022-02-21 [1] CRAN (R 4.1.2)
+# ggthemes        * 4.2.4   2021-01-20 [1] CRAN (R 4.1.2)
+# pals            * 1.7     2021-04-17 [1] CRAN (R 4.1.0)
+# Polychrome      * 1.3.1   2021-07-16 [1] CRAN (R 4.1.2)
+# rcartocolor     * 2.0.0   2019-05-03 [1] CRAN (R 4.1.0)
+# RColorBrewer    * 1.1-2   2014-12-07 [1] CRAN (R 4.1.0)
+# reticulate      * 1.24    2022-01-26 [1] CRAN (R 4.1.2)
+# scico           * 1.3.0   2021-12-08 [1] CRAN (R 4.1.2)
+# shiny           * 1.7.1   2021-10-02 [1] CRAN (R 4.1.2)
+# viridisLite     * 0.4.0   2021-04-13 [1] CRAN (R 4.1.0)
 
-# library(paletteer)
-# https://github.com/EmilHvitfeldt/r-color-palettes
 
-c4a_series_remove(are.you.sure = TRUE)
+
+
+
+c4a_palettes_remove(are.you.sure = TRUE)
 
 ###################################
 ### package grDevices: cat
@@ -27,51 +39,23 @@ local({
 
 	p1 = pals[c("R3", "R4", "ggplot2", "Okabe-Ito")]
 	names(p1) = c("R3", "R4", "ggplot2", "okabe")
-	s1 = "misc"
 
-	# p2 = pals[c("Accent", "Dark 2", "Paired", "Pastel 1", "Pastel 2", "Set 1", "Set 2", "Set 3")]
-	# s2 = "brewer"
-
-	#p3 = pals[c("Alphabet", "Polychrome 36")]
-	#s3 = "misc"
-
-	c4a_series_add(p1, types = "cat", series = s1)
-	#c4a_series_add(p2, types = "cat", series = s2c4a_series_add)
-	#c4a_series_add(p3, types = "cat", series = s3, take.gray.for.NA = FALSE, remove.other.grays = FALSE, remove.blacks = FALSE)
-	invisible(NULL)
+	c4a_palettes_add(p1, types = "cat", series = "misc")
 })
 
[…]

### 2022-03-29 examples/c4a_palettes_add.R

@@ -1,37 +1,35 @@
 #######################################
 # Palettes from Statistics Netherlands
 #######################################
+
 cbsnl_cols = list(
 	charts = c("#00a1cd", "#0058b8", "#afcb05", "#53a31d", "#d9328a", "#7d4791",
 			   "#f39200", "#c90c0f", "#0581a2", "#163a72", "#899d0c", "#488225",
 			   "#af0e80", "#56217a", "#da5914", "#9c1006"),
-	map_blue7 = c("#e1f4fd", "#c0e7ff", "#77cbe5", "#3d95d4", "#2256a0", "#143564",
-				  "#09183c"),
-	map_green7 = c("#f1f6de", "#edf0c7", "#c9de85", "#85bc22", "#348a3a", "#0f5f34",
-				   "#114625"),
-	map_red7 = c("#fedfc7", "#ffc597", "#f89e6b", "#e74d15", "#c01f26", "#82001e",
-				 "#5b0708"),
-	map_purple7 = c("#fbe2ed", "#f8c1d9", "#e38cbf", "#be3e8d", "#8b176f", "#490045",
-					"#2d002c"),
-	map_blue5 = c("#c0e7ff", "#77cbe5", "#3d95d4", "#2256a0", "#143564"),
-	map_green5 = c("#edf0c7", "#c9de85", "#85bc22", "#348a3a", "#0f5f34"),
-	map_red5 = c("#ffc597", "#f89e6b", "#e74d15", "#c01f26", "#82001e"),
-	map_purple5 = c("#f8c1d9", "#e38cbf", "#be3e8d", "#8b176f", "#490045"),
+	map_blue = structure(c("#e1f4fd", "#c0e7ff", "#77cbe5", "#3d95d4", "#2256a0", "#143564",
+				  "#09183c"), index = list('5' = 2:6, '6' = 2:7, '7' = 1:7)),
+	map_green = structure(c("#f1f6de", "#edf0c7", "#c9de85", "#85bc22", "#348a3a", "#0f5f34",
+				   "#114625"), index = list('5' = 2:6, '6' = 2:7, '7' = 1:7)),
+	map_red = structure(c("#fedfc7", "#ffc597", "#f89e6b", "#e74d15", "#c01f26", "#82001e",
+				 "#5b0708"), index = list('5' = 2:6, '6' = 2:7, '7' = 1:7)),
[…]

### 2022-03-31 README.Rmd
c4a_citation added

@@
 * `c4a_palettes` Get available palette names
-* `c4a_series` Get available series
-* `c4a_meta` Get meta information (such as type and maximum number of colors )
-* `c4a_ls` Environment via which palette names can be browsed with auto-completion (using `$`)
+* `c4a_series` Get available series names
+* `c4a_info` Get palette information (such as type and maximum number of colors)
+* `c4a_citation` Show how to cite palettes (with bibtex code).
+* `.P` Environment via which palette names can be browsed with auto-completion (using `$`)
 

### 2022-03-31 R/c4a.R

@@ -36,7 +36,7 @@ c4a = function(palette = NULL, n = NA, m = NA, type = c("cat", "seq", "div", "bi
 		mes = NULL
 	}
 
-	x = c4a_meta(palette, verbose = verbose)
+	x = c4a_info(palette, verbose = verbose)
 
 	reverse = xor(reverse, x$reverse)
 
@@ -78,16 +78,16 @@ c4a = function(palette = NULL, n = NA, m = NA, type = c("cat", "seq", "div", "bi
 	}
 }
 
-#' Get meta information from a cols4all palette
+#' Get information from a cols4all palette
 #'
-#' Get meta information from a cols4all palette
+#' Get information from a cols4all palette
 #'
 #' @param palette name of the palette
 #' @param no.match what happens is no match is found? Options: `"error"`: an error is thrown, `"null"`: `NULL` is returned
 #' @param verbose should messages be printed?
 #' @return list with the following items: name, series, fullname, type, palette (colors), na (color), nmax, and reverse. The latter is `TRUE` when there is a `"-"` prefix before the palette name.
 #' @export
-c4a_meta = function(palette, no.match = c("error", "null"), verbose = TRUE) {
+c4a_info = function(palette, no.match = c("error", "null"), verbose = TRUE) {
 	isrev = (substr(palette, 1, 1) == "-")
 	if (isrev) palette = substr(palette, 2, nchar(palette))
 
@@ -121,7 +121,7 @@ c4a_na = function(palette = NULL, type = c("cat", "seq", "div"), verbose = TRUE)
 	}
 
 
-	x = c4a_meta(palette, verbose = verbose)
+	x = c4a_info(palette, verbose = verbose)
 
 	if (!is.null(mes) && verbose) message(mes)
 

### 2022-03-31 R/c4a_citation.R

@@ -1,0 +1,67 @@
+#' Show how to cite palettes
+#'
+#' Show how to cite palettes
+#'
+#' @param name name of a palette or series
+#' @export
+c4a_citation = function(name) {
+	is_series = name %in% c4a_series()
+
+	if (!is_series) {
+		inf = c4a_info(name, no.match = NULL, verbose = FALSE)
+		if (is.null(inf)) {
+			message(paste0("\"", name, "\" is not a known series or palette name"))
+			return(invisible(NULL))
+		} else {
+			name = inf$name
+			series = inf$series
+		}
+	} else {
+		series = name
+	}
+
+	bnames = names(.C4A$zbib)
+
+	if (is_series) {
+		if (!(name %in% bnames)) {
+			message("No citation information (bibentry) found for series \"", series, "\"")
+			return(invisible(NULL))
+		}
+		bib = .C4A$zbib[[name]]
+	} else {
+		split_names = lapply(bnames, function(bn) {
+			strsplit(bn, split = ".", fixed = TRUE)[[1]][-1]
+		})
+		split_series = sapply(bnames, function(bn) {
+			strsplit(bn, split = ".", fixed = TRUE)[[1]][1]
+		})
+		match_name = sapply(split_names, function(sn) {
+			name %in% sn
+		})
+		match_series = (series == split_series)
+		if (any(match_name & match_series)) {
+			bib = .C4A$zbib[[which(match_name & match_series)]]
+		} else {
+			if (!(series %in% bnames)) {
+				message("No citation information (bibentry) found for palette \"", name, "\" from series \"", series, "\"")
+				return(invisible(NULL))
+			}
+			is_series = TRUE
+			bib = .C4A$zbib[[series]]
+		}
+	}
+
+	cat("\n")
+
+	if (is_series) {
+		cat("To cite palettes from", series, "in publications use:\n\n")
+	} else {
+		cat("To cite palette", name, "from", series, "in publications use:\n\n")
+	}
+
+	print(bib, style = "text")
+	cat("\n")
+
+	print(bib, style = "Bibtex")
+	invisible(bib)
+}

### 2022-03-31 R/onLoad.R

@@ -13,8 +13,9 @@
 #' @section Palette names and properties:
 #' \tabular{ll}{
 #' \code{\link{c4a_palettes}}\tab Get available palette names \cr
-#' \code{\link{c4a_series}}\tab Get available series \cr
-#' \code{\link{c4a_meta}}\tab Get meta information (such as type and maximum number of colors) \cr
+#' \code{\link{c4a_series}}\tab Get available series names\cr
+#' \code{\link{c4a_info}}\tab Get palette information (such as type and maximum number of colors) \cr
+#' \code{\link{c4a_citation}}\tab Show how to cites palettes (with bibtex code) \cr
 #' \code{\link{.P}}\tab Environment via which palette names can be browsed with auto-completion (using `$`) \cr
 #' }
 #'
@@ -25,95 +26,6 @@
 #' \code{\link{c4a_sysdata_import}}\tab Import system data \cr
 #' \code{\link{c4a_sysdata_export}}\tab Export system data \cr
 #' }
-#'
-#' @section References:
-#'
-#' \strong{misc}
-#'
-#' * `r3` and `r4` base R colors, from the `grDevices` package
-#' * `ggplot2` from the `ggplot2` package (also included in `grDevices`)
-#' * `okabe` by Okabe-Ito (2002, revised 2008), Color Universal Design (CUD) - How to make figures and presentations that are friendly to Colorblind people -, \url{https://jfly.uni-koeln.de/color/} (also included in `grDevices`)
-#' * `alphabet` from the `Polychroma` package (also included in `grDevices`)
-#' * `alphabet2` by P.Green-Armytage (2010): A Colour Alphabet and the Limits of Colour Coding. Colour: Design & Creativity (5) (2010): 10, 1-23. www.aic-color.org/journal/v5/jaic_v5_06.pdf
-#' * `polychrome36` from the `Polychrome` package (also included in `grDevices`)
-#' * `kelly` by K. Kelly (1965): Twenty-two colors of maximum contrast. Color Eng., 3(6), 1965. http://www.iscc.org/pdf/PC54_1724_001.pdf (obtained from the `pals` package)
-#' * `watlington` by John Watlington. \url{http://alumni.media.mit.edu/~wad/color/palette.html} (obtained from the `pals` package)
[…]

### 2022-03-31 build/build_palettes.R

@@ -581,5 +581,10 @@ local({
 .s = get("s", .C4A)
 
 
-save(.z, .s, file="R/sysdata.rda", compress="xz")
+
+.zbib = bibtex::read.bib("build/references.bib")
+
+
+save(.z, .s, .zbib, file="R/sysdata.rda", compress="xz")
+
 

### 2022-04-03 README.Rmd
should be pretty stable: to CRAN?

@@
 :-------------------------:|:-------------------------:|:-------------------------:
-[![](https://user-images.githubusercontent.com/2444081/156847689-7e1a4608-9b03-45f5-9ccf-56e6d8844d47.png)](https://user-images.githubusercontent.com/2444081/156847689-7e1a4608-9b03-45f5-9ccf-56e6d8844d47.png) | [![](https://user-images.githubusercontent.com/2444081/156847703-b429c521-89d5-4e14-85c1-0f100ca93b45.png)](https://user-images.githubusercontent.com/2444081/156847703-b429c521-89d5-4e14-85c1-0f100ca93b45.png) | [![](https://user-images.githubusercontent.com/2444081/156847712-33a3cd22-a50b-44e2-9666-b67fcb0d2853.png)]((https://user-images.githubusercontent.com/2444081/156847712-33a3cd22-a50b-44e2-9666-b67fcb0d2853.png))
+[![](https://user-images.githubusercontent.com/2444081/161420545-68c007d1-640e-47f1-84e3-1e22c7b74e20.png)](https://user-images.githubusercontent.com/2444081/161420545-68c007d1-640e-47f1-84e3-1e22c7b74e20.png) | [![](https://user-images.githubusercontent.com/2444081/161420541-1c2dfefc-0451-47ed-badd-d5ecc339ead0.png)](https://user-images.githubusercontent.com/2444081/161420541-1c2dfefc-0451-47ed-badd-d5ecc339ead0.png) | [![](https://user-images.githubusercontent.com/2444081/161420543-62bc6947-df35-43e2-8277-35b2dbc223ca.png)]((https://user-images.githubusercontent.com/2444081/161420543-62bc6947-df35-43e2-8277-35b2dbc223ca.png)) |
+
+Bivariate (seq x seq)		|  Bivariate (seq x cat)	|		Bivariate (seq x desaturated)
+:-------------------------:|:-------------------------:|:-------------------------:
+[![](https://user-images.githubusercontent.com/2444081/161420551-84475286-5394-4379-abfc-d765567c2db6.png)]((https://user-images.githubusercontent.com/2444081/161420551-84475286-5394-4379-abfc-d765567c2db6.png)) | [![](https://user-images.githubusercontent.com/2444081/161420550-dc0ac110-ddea-4f02-8d47-cbb6221cca4f.png)]((https://user-images.githubusercontent.com/2444081/161420550-dc0ac110-ddea-4f02-8d47-cbb6221cca4f.png)) | [![](https://user-images.githubusercontent.com/2444081/161420549-8130172f-9047-48c3-b8f5-1a3e7ae7c3b3.png)]((https://user-images.githubusercontent.com/2444081/161420549-8130172f-9047-48c3-b8f5-1a3e7ae7c3b3.png))
 
@@
 Finally, for each color palette a color for missing values is assigned, which is especially important for spatial data visualization.
-Currently we support several types: *categorical* (qualitative) palettes, *sequential* palettes, *diverging* palettes, and *bivariate* palettes (divided into three types).
-In the near future, more palette types may be added, such as *cyclic* and *hierarchical*.
+Currently we support several types: *categorical* (qualitative) palettes, *sequential* palettes, *diverging* palettes, and *bivariate* palettes (divided into three subtypes).
 
@@
 # install.packages("devtools")
-devtools::install_github("mtennekes/cols4all")
+remotes::install_github("mtennekes/cols4all")
 ```
 
-```{r, echo=FALSE}
-getwd()
+```{r, echo=FALSE, message=FALSE}
 devtools::load_all(".")
@@
 library(cols4all)
-# For the GUI, shiny, shinyjs, and kableExtra are required
-# install.packages(c("shiny", "shinyjs", "kableExtra"))
 c4a_gui()
 ```
-
-Besides browsing through palettes, it is also easy to copy color codes to the clipboard, either by selecting the hidden text in the color tables or by clicking on the icons at the right-hand-side:
-
-![](https://user-images.githubusercontent.com/2444081/156849917-59af1dc1-ee5e-40be-bf87-5c7384351ae0.png){width=50%}
 
@@
 
-* It will be possible to submit an own series of color palettes.
+* It is be possible to submit own color palettes.
 
@@
 
-* Do you miss certain palettes or series?
+* Do you miss certain palettes?
 

### 2022-04-03 R/attach_bib.R

@@ -1,0 +1,37 @@
+attach_bib = function() {
+	z = .C4A$z
+	zbib = .C4A$zbib
+
+	bnames = names(zbib)
+	split_names = lapply(bnames, function(bn) {
+		strsplit(bn, split = ".", fixed = TRUE)[[1]][-1]
+	})
+	split_series = sapply(bnames, function(bn) {
+		strsplit(bn, split = ".", fixed = TRUE)[[1]][1]
+	})
+
+
+	nms = mapply(function(x,y) {
+		if (length(x)) paste(y, x, sep = ".") else y
+	}, split_names, split_series, SIMPLIFY = FALSE)
+
+	match(z$series, split_series)
+
+	ids = mapply(function(fn, s) {
+		id = which(sapply(nms, function(nm) {
+			any(fn %in% nm | s %in% nm)
+		}))
+		if (!length(id)) id = 0
+		id
+	}, z$fullname, z$series, SIMPLIFY = TRUE, USE.NAMES = FALSE)
+
+	zbib_cit = sapply(zbib, format, style = "text", USE.NAMES = FALSE)
+	zbib_bib = sapply(zbib, format, style = "Bibtex", USE.NAMES = FALSE)
+
+	ids[ids==0] = NA
+
+	z$cit = zbib_cit[ids]
+	z$bib = zbib_bib[ids]
+	.C4A$z = z
+	NULL
+}

### 2022-04-03 R/c4a_gui.R

@@ -246,7 +246,7 @@ c4a_gui = function(type = "cat", n = NA, series = c("misc", "brewer", "hcl", "to
 			shiny::showModal(shiny::modalDialog(title = "Number of palettes per series (rows) and type (columns)",
 												shiny::renderTable(tab_k, na = "", striped = TRUE, hover = TRUE, bordered = TRUE),
 												shiny::div(style="font-size: 75%;", shiny::renderTable(.C4A$type_info)),
-												footer = modalButton("Close"),
+												footer = shiny::modalButton("Close"),
 												style = "color: #000000;"))
 		})
 

### 2022-04-03 R/c4a_overview.R

@@ -1,17 +1,0 @@
-c4a_overview = function(type) {
-	z = .C4A$z
-
-	zt = z[z$type == type, ]
-
-	zt$series = as.factor(zt$series)
-
-	np = tapply(zt$name, INDEX = zt$series, FUN = length)
-	nmin = tapply(zt$nmin, INDEX = zt$series, FUN = min)
-	nmax = tapply(zt$nmax, INDEX = zt$series, FUN = max)
-
-	df = data.frame(Series = levels(zt$series), np = np, nmin = nmin, nmax = nmax)
-	rownames(df) = NULL
-
-	kableExtra::kbl(df, col.names = c("Series", "Number of palettes", "Min. number of colors", "Max. number of colors")) |>
-		kableExtra::kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
-}

### 2022-04-03 R/c4a_palettes_add.R

@@ -160,6 +160,8 @@ c4a_palettes_add = function(x, xNA = NA, types, series, nmin = NA, nmax = NA, nd
 
 
 	if (!is.null(.z)) {
+		.z$bib = NULL
+		.z$cit = NULL
 		if (any(fnms %in% .z$fullname)) stop("Fulnames already exist: ", paste(intersect(fnms, .z$fullname), collapse = ", "))
 
 		z = rbind(.z, z)
@@ -171,6 +173,7 @@ c4a_palettes_add = function(x, xNA = NA, types, series, nmin = NA, nmax = NA, nd
 	.C4A$z = z
 	.C4A$zbib = zbib
 	.C4A$s = s
+	attach_bib()
 	fill_P()
 	invisible(NULL)
 }

### 2022-04-03 R/c4a_sysdata_import.R

@@ -9,16 +9,18 @@
 #' @name c4a_sysdata_import
 #' @export
 c4a_sysdata_import = function(x) {
-	if (!is.list(x) || !setequal(c("z", "s"), names(x))) stop("x should be a list of two: z and s", call. = FALSE)
+	if (!is.list(x) || !setequal(c("z", "s", "zbib"), names(x))) stop("x should be a list of three: z, s, zbib", call. = FALSE)
 
 	z = check_z(x$z)
 	s = check_s(x$s, nrow(z))
-
+	zbib = x$zbib # to do: check
 
 	message("cols4all system data imported successfully")
 
 	.C4A$z = z
 	.C4A$s = s
+	.C4A$zbib = zbib
+	attach_bib()
 	fill_P()
 	invisible(NULL)
 }
@@ -27,6 +29,10 @@ c4a_sysdata_import = function(x) {
 #' @name c4a_sysdata_export
 #' @export
 c4a_sysdata_export = function() {
-	list(z = .C4A$z,
-		 s = .C4A$s)
+	z = .C4A$z
+	z$cit = NULL
+	z$bib = NULL
+	list(z = z,
+		 s = .C4A$s,
+		 zbib = .C4A$zbib)
 }

### 2022-04-03 R/check_pals.R

@@ -47,6 +47,8 @@ check_bivs_pal = function(p) {
 	p2 = p[,1]
 	pd = diag(p)
 
+	#if (nrow(p) == 5) browser()
+
 	x12 = check_div_pal(c(rev(p1[-1]), p2))
 	x1d = check_div_pal(c(rev(p1[-1]), pd))
 	x2d = check_div_pal(c(rev(p2[-1]), pd))

### 2022-04-03 R/create_biv_palette.R

@@ -1,3 +1,13 @@
+aregreys = function(x) {
+	all(get_hcl_matrix(x)[,2] < 1)
+}
+
+# ensure the diagonal are pure greys
+convert2grey = function(x) {
+	greys = rowSums(attr(hex2RGB(x), "coords")) / 3
+	rgb(greys, greys, greys)
+}
+
 create_biv_palette = function(palette, biv.method) {
 	if (!(biv.method %in% c("div2seqseq", "div2catseq", "seq2uncseq", "byrow", "bycol"))) {
 		n = as.integer(substr(biv.method, nchar(biv.method), nchar(biv.method)))
@@ -26,13 +36,14 @@ create_biv_palette = function(palette, biv.method) {
 	if (biv.method == "div2seqseq") {
 		a = get_hcl_matrix(palette)
 
-
+		# hue matrix
 		h1 = matrix(a[n:1, 1], nrow = n, ncol = n, byrow = FALSE)
 		h2 = matrix(a[n:np, 1], nrow = n, ncol = n, byrow = TRUE)
 		h = matrix(a[n, 1], nrow = n, ncol = n)
 		h[lower.tri(h)] = h1[lower.tri(h)]
 		h[upper.tri(h)] = h2[upper.tri(h)]
 
+		# chroma matrix
 		cr1 = a[n:1, 2]
 		cr2 = a[n:np, 2]
 
@@ -42,7 +53,7 @@ create_biv_palette = function(palette, biv.method) {
 			cr[1:i, i] = seq(cr2[i], 0, length.out = i)
 		}
 
-
+		# limunance matrix
 		la1 = a[n:1, 3]
 		la2 = a[n:np, 3]
 		la0 = (la1 + la2) / 2
@@ -54,9 +65,7 @@ create_biv_palette = function(palette, biv.method) {
 			xx <- seq(1,n*sqrt(2),length.out=n)
 			stats::predict(fit2, data.frame(x=xx))
 		})
-
 		l = matrix(la0[1], nrow = n, ncol = n)
-
 		for (i in 2:n) {
 			l[i, 1:i] = seq(la1[i], la0b[i], length.out = i)
 			l[1:i, i] = seq(la2[i], la0b[i], length.out = i)
@@ -71,7 +80,9 @@ create_biv_palette = function(palette, biv.method) {
 		# l[lower.tri(l)] = l1[lower.tri(l1)]
[…]

### 2022-04-03 R/onLoad.R

@@ -46,7 +46,7 @@ NULL
 	assign("s", .s, envir = .C4A)
 	assign("zbib", .zbib, envir = .C4A)
 
-
+	attach_bib()
 
 	with(.C4A,{
 		defaults = c(cat = "tol.muted", seq = "hcl.blues2", div = "hcl.purple-green")

### 2022-04-03 build/build_palettes.R

@@ -528,35 +528,78 @@ local({
 
 
 
-
 ########################################################################################
 ######################################## BIVARIATE #####################################
 ########################################################################################
-
+c4a_palettes_remove(series = "c4a")
 local({
-	#c4a_palettes(type = "div")
-	pals = c("hcl.purple_brown", "hcl.purple_green")
+	bu2 = c4a("hcl.blues3", n = 5, range = c(0.3, 0.8))
+	yl_rd = c4a("hcl.yellow_red", n = 5, range = c(0.3, 0.8))
+	pg = hcl.colors(11, "Purple-Green")
+	bu = hcl.colors(9, "Blues 3")[7:3]
+	gn = hcl.colors(9, "Greens 3")[7:3]
+	pu = hcl.colors(9, "Purples 3")[7:3]
 
-	b = .C4A$z[match(pals, .C4A$z$fullname), ]
-	p = b$palette
-	names(p) = c("pu_br_divs", "pu_gn_divs")
+	# extract average c and l values for both wings
+	mat = get_hcl_matrix(pg)
+	cs = (mat[1:6, 2] + mat[11:6, 2]) / 2
+	cs[6] = 0
+	cs[5] = 25
+	ls = (mat[1:6, 3] + mat[11:6, 3]) / 2
 
-	p = lapply(p, function(pal)pal[3:9])
+	ls2 = ls
+	ls2[6] = 92
 
-	c4a_palettes_add(p, types = "bivs", series = "c4a", biv.method = "div2seqseq", space = "Lab")
 
-	p2 = b$palette
-	names(p2) = c("pu_br_divc", "pu_gn_divc")
+	# candidate hues
+	hs = seq(0, 359, by = 0.2)
 
-	p2 = lapply(p2, function(pal)pal[2:10])
-	c4a_palettes_add(p2, types = "bivc", series = "c4a", biv.method = "div2catseq", space = "Lab")
+	# find hues for which colors have most chroma (the higher the better to distinguisch with grey)
+	res = lapply(1:5, function(i) {
+		sapply(hs, function(h) hcl(h=h,c=cs[i], l =ls[i]))
+	})
[…]

### 2022-11-17 R/c4a_gui.R
getting ready for CRAN

@@ -120,13 +120,13 @@ c4a_gui = function(type = "cat", n = NA, series = "all") {
 			right = 40,
 			width = 90,
 			shiny::checkboxInput("dark", "Dark mode", value = FALSE)),
-		shiny::absolutePanel(
-			top = 10,
-			right = 10,
-			width = 20,
-			shiny::actionButton("info", "", shiny::icon("info"),
-								style="color: #fff; background-color: #337ab7; border-color: #2e6da4")
-		),
+		# shiny::absolutePanel(
+		# 	top = 10,
+		# 	right = 10,
+		# 	width = 20,
+		# 	shiny::actionButton("info", "", shiny::icon("info"),
+		# 						style="color: #fff; background-color: #337ab7; border-color: #2e6da4")
+		# ),
 		shiny::tabsetPanel(
 			id="inTabset",
 			shiny::tabPanel("Overview",
@@ -384,12 +384,14 @@ c4a_gui = function(type = "cat", n = NA, series = "all") {
 
 
 		tab_vals = shiny::reactiveValues(pal = pal_init,
+										 na = FALSE,
 										 palBW = c(pal_init, "#FFFFFF", "#000000"),
 										 pal_name = palette,
 										 n = n_init,
 										 colA1 = pal_init[1], colA2 = pal_init[2],
 										 colB1 = pal_init[1], colB2 = pal_init[2],
 										 colC1 = pal_init[1], colC2 = pal_init[2],
+										 CR = colorspace::contrast_ratio(pal_init[1], pal_init[2]),
 										 type = type12,
 										 cvd = "none",
 										 b = approx_blues(pal_init),
@@ -585,11 +587,13 @@ c4a_gui = function(type = "cat", n = NA, series = "all") {
 			if (length(pals)) {
 				tab_vals$pals = pals
 				tab_vals$n = n
-				tab_vals$pals = pals
 				if (!length(tab_vals$pal_name) || !(tab_vals$pal_name %in% pals)) {
 					tab_vals$pal_name = pals[1]
 				}
 				cols = as.vector(c4a(tab_vals$pal_name, n = n))
+				na = values$na
+				tab_vals$na = na
+				if (na) cols = c(cols, c4a_na(tab_vals$pal_name))
 				tab_vals$pal = cols
 				tab_vals$palBW = c(cols, "#FFFFFF", "#000000")
 				tab_vals$type = values$type
@@ -597,6 +601,7 @@ c4a_gui = function(type = "cat", n = NA, series = "all") {
[…]

### 2022-11-17 R/c4a_name_convert.R

@@ -35,8 +35,6 @@ check_name_presence = function(name, verbose = TRUE) {
 #' @param name palette name to be converted
 #' @param no.match what happens is no match is found? Options: `"error"`: an error is thrown, `"null"`: `NULL` is returned
 #' @param verbose should messages be printed?
-#' @example ./examples/c4a_name_convert.R
-#' @export
 c4a_name_convert = function(name, no.match = c("error", "null"), verbose = TRUE) {
 	no.match = match.arg(no.match)
 	name2 = check_name_presence(name, verbose)

### 2022-11-17 R/c4a_palettes.R

@@ -1,6 +1,6 @@
 #' Get available palette names and series
 #'
-#' `c4a_palettes` lists all available cols4all color palettes. Palettes are organized by series. The available series are listed with `c4a_series`. Palettes are also organized per functional type, where we currently support: categorical `"cat"`, sequential `"seq"`, and diverging `"div"`" palette types. The function `c4a_default_palette` returns the default (recommended) palette per type. In an IDE with auto-completion (such as RStudio) it is possible to browse through the palette names with `.P` (using `$` like in lists).
+#' `c4a_palettes` lists all available cols4all color palettes. Palettes are organized by series. The available series are listed with `c4a_series`. Palettes are also organized per functional type, where we currently support: categorical `"cat"`, sequential `"seq"`, and diverging `"div"`" palette types. In an IDE with auto-completion (such as RStudio) it is possible to browse through the palette names with `.P` (using `$` like in lists).
 #'
 #' @param type type of color palette: one of `"all"` (all palettes), `"cat"` (categorical/qualitative palettes), `"seq"` (sequential palettes) and `"div"` (diverging palettes).
 #' @param series series to list the palettes from. Run `c4a_series` to see the options.
@@ -30,10 +30,7 @@ c4a_series = function(type = c("all", "cat", "seq", "div")) {
 	unique({if (type != "all") series[z$type == type] else series})
 }
 
-#' @rdname c4a_palettes
-#' @name c4a_default_palette
-#' @export
-c4a_default_palette = function(type = c("cat", "seq", "div")) {
+c4a_default_palette = function(type) {
 	type = match.arg(type)
 	.C4A$defaults[type]
 }

### 2022-11-17 R/c4a_plot_X_matrix.R

@@ -161,8 +161,8 @@ plot_matrix = function(p, id1 = NULL, id2 = NULL, type = c("CR", "dist"), cvd = 
 
 symbol_size = function(cr, type) {
 	brks = .C4A$matrix_breaks[[type]]
-	brks_digits = .C4A$matrix_breaks_digits[type]
-	cr = round(cr, brks_digits)
+	#brks_digits = .C4A$matrix_breaks_digits[type]
+	#cr = round(cr, brks_digits)
 	id = which(cr <= brks[-1])[1]
 	if (is.na(id)) id = length(brks)
 	id

### 2022-11-17 R/c4a_plot_map.R

@@ -1,4 +1,4 @@
-c4a_plot_map = function(cols = NULL, col1 = "blue", col2 = "red", borders = "black", lwd = 0, crop = FALSE, dark = FALSE, dist = c("random", "gradient")) {
+c4a_plot_map = function(cols = NULL, col1 = "blue", col2 = "red", borders = "black", lwd = 0, include.na = FALSE, crop = FALSE, dark = FALSE, dist = c("random", "gradient")) {
 	dist = match.arg(dist)
 
 
@@ -8,14 +8,16 @@ c4a_plot_map = function(cols = NULL, col1 = "blue", col2 = "red", borders = "bla
 		cols = c(col1, col2)
 	}
 
-	n = length(cols)
+	n = length(cols) - include.na
 	k = length(shp$gp$fill)
 
 
 	shp$gp$fill = if (dist == "random") {
 		rep(cols, length.out = k)
 	} else {
-		cols[round(seq(1 - 0.5/n, n + 0.5/n, length.out = k))]#[order(shp_c[, 1])]
+		x = cols[round(seq(1 - 0.5/n, n + 0.5/n, length.out = k))]#[order(shp_c[, 1])]
+		if (include.na) x[sample.int(k, round(k/10))] = cols[n+1]
+		x
 	}
 
 

### 2022-11-17 R/c4a_table.R

@@ -33,7 +33,7 @@ table_columns = function(type, show.scores) {
 #' Graphical user interface to analyse palettes. `c4a_table` shows a table that can be opened in the browser. `c4a_gui` is a graphical user interface (shiny app) around this table.
 #'
 #' @param type type of palette: `"cat"` for categorical (aka qualitative), `"seq"` for sequential, `"div"` for diverging, and `"bivs"`/`"bivc"`/`"bivd"`/`"bivg"` for bivariate (seq-seq, seq-cat, seq-div, and uncertainty-seq). For `c4a_gui` it only determines which type is shown initially.
-#' @param n,m for univariate palettes, `n` is the number of displayed colors. For bivariate palettes `"biv"`, `n` and `m` are the number of columns and rows respectively. If omitted: for `"cat"` the full palette is displayed, for `"seq"` and `"div"`, 9 colors, and for `"bivs"`/`"bivc"`/`"bivd"`/`"bivg"` 4 columns and rows. For `c4a_gui` it only determines which number of colors initially.
+#' @param n,m `n` is the number of displayed colors. For bivariate palettes `"biv"`, `n` and `m` are the number of columns and rows respectively. If omitted: for `"cat"` the full palette is displayed, for `"seq"` and `"div"`, 9 colors, and for `"bivs"`/`"bivc"`/`"bivd"`/`"bivg"` 4 columns and rows. For `c4a_gui` it only determines which number of colors initially.
 #' @param cvd.sim color vision deficiency simulation: one of `"none"`, `"deutan"`, `"protan"`, `"tritan"`
 #' @param sort column name to sort the data. For column names, see details. Use a `"-"` prefix to reverse the order.
 #' @param text.format The format of the text of the colors. One of `"hex"`, `"RGB"` or `"HCL"`.

### 2023-01-05 README.Rmd
updated readme, fuzzy name matching,

--- 2022-03-29 README.Rmd
@@
 output: github_document
+editor_options: 
+  chunk_output_type: console
 ---
@@
 )
+hook_output <- knitr::knit_hooks$get("output")
+knitr::knit_hooks$set(output = function(x, options) {
+   lines <- options$output.lines
+   if (is.null(lines)) {
+     return(hook_output(x, options))  # pass to default hook
+   }
+   x <- unlist(strsplit(x, "\n"))
+   more <- "..."
+   if (length(lines)==1) {        # first n lines
+     if (length(x) > lines) {
+       # truncate the output, but add ....
+       x <- c(head(x, lines), more)
+     }
+   } else {
+     x <- c(more, x[lines], more)
+   }
+   # paste these lines together
+   x <- paste(c(x, ""), collapse = "\n")
+   hook_output(x, options)
+ })
 ```
@@
 
-Categorical             |  Sequential | Diverging
-:-------------------------:|:-------------------------:|:-------------------------:
-[![](https://user-images.githubusercontent.com/2444081/156847689-7e1a4608-9b03-45f5-9ccf-56e6d8844d47.png)](https://user-images.githubusercontent.com/2444081/156847689-7e1a4608-9b03-45f5-9ccf-56e6d8844d47.png) | [![](https://user-images.githubusercontent.com/2444081/156847703-b429c521-89d5-4e14-85c1-0f100ca93b45.png)](https://user-images.githubusercontent.com/2444081/156847703-b429c521-89d5-4e14-85c1-0f100ca93b45.png) | [![](https://user-images.githubusercontent.com/2444081/156847712-33a3cd22-a50b-44e2-9666-b67fcb0d2853.png)]((https://user-images.githubusercontent.com/2444081/156847712-33a3cd22-a50b-44e2-9666-b67fcb0d2853.png))
-
 Color palettes are well organized and made consistent with each other.
@@
 Finally, for each color palette a color for missing values is assigned, which is especially important for spatial data visualization.
-Currently we support several types: *categorical* (qualitative) palettes, *sequential* palettes, *diverging* palettes, and *bivariate* palettes (divided into three types).
-In the near future, more palette types may be added, such as *cyclic* and *hierarchical*.
+Currently we support several types: *categorical* (qualitative) palettes, *sequential* palettes, *diverging* palettes, and *bivariate* palettes (divided into four subtypes).
 
@@
 
-**cols4all** will be available on CRAN soon. Until then it can be installed using:
+**cols4all** will be available on CRAN in January or February this year (2023). Until then it can be installed using:
 
 ```{r, eval=FALSE}
-# install.packages("devtools")
-devtools::install_github("mtennekes/cols4all")
+install.packages("remotes")
+remotes::install_github("mtennekes/cols4all", dependencies = TRUE)
 ```
 
-```{r, echo=FALSE}
-getwd()
+```{r, echo=FALSE, message=FALSE}
 devtools::load_all(".")
@@
 
-Navigation through the color palettes is easy via the GUI shown above, which is started with:
+Load the package:
 
@@
 library(cols4all)
-# For the GUI, shiny, shinyjs, and kableExtra are required
-# install.packages(c("shiny", "shinyjs", "kableExtra"))
+```
+
+The main tool is a dashboard, which is started with:
+
+```{r, eval=FALSE}
 c4a_gui()
@@
 
-Besides browsing through palettes, it is also easy to copy color codes to the clipboard, either by selecting the hidden text in the color tables or by clicking on the icons at the right-hand-side:
+![https://user-images.githubusercontent.com/2444081/210850914-cdb8a128-1b8a-4900-94d0-dfa6ca449585.png](https://user-images.githubusercontent.com/2444081/210850914-cdb8a128-1b8a-4900-94d0-dfa6ca449585.png)
 
-![](https://user-images.githubusercontent.com/2444081/156849917-59af1dc1-ee5e-40be-bf87-5c7384351ae0.png){width=50%}
+What palettes are available?
 
-Selecting a palette is easy:
+```{r, output.lines = 10}
+# All loaded palettes
+c4a_palettes()
+```
+
+```{r}
+# Diverging palettes from the 'hcl' series
+c4a_palettes(type = "div", series = "hcl")
+```
+
+Give me the colors!
 
@@
 c4a("kelly", 7)
-
-# find names of hcl palettes that are diverging
-c4a_palettes(type = "div", series = "hcl")
 
@@
 
-* `c4a_gui` GUI (shiny app) to see and analyse the palettes
-* `c4a` Get the colors of a palette
+*   `c4a_gui` Dashboard for analyzing the palettes
+*   `c4a` Get the colors from a palette
 
@@
 *   `c4a_palettes` Get available palette names
-* `c4a_series` Get available series
-* `c4a_meta` Get meta information (such as type and maximum number of colors )
-* `c4a_ls` Environment via which palette names can be browsed with auto-completion (using `$`)
+*   `c4a_series` Get available series names
+*   `c4a_citation` Show how to cite palettes (with bibtex code).
+*   `c4a_info` Get information from a palette, such as type and maximum number
+    of colors
+*   `.P` Environment via which palette names can be browsed with
+    auto-completion (using `$`)
 
@@
 
-* `c4a_palettes_add` Add color palettes
-* `c4a_palettes_remove` Remove color palettes
+*   `c4a_data` Build color palette data
+*   `c4a_load` Load color palette data
 *   `c4a_sysdata_import` Import system data
@@
 
-* Colors for missing values are made explicit.
-For this, either a greyscale color that is already contained in the palette is used, or a greyscale color is selected that is distinguisable, also for people with color vision deficiency.
+*	Color palettes are characterized and analysed. Properties such as color blindness, fairness (whether colors stand out about equally), and contrast are determined for each palette.
 
-* Palettes are made consistent with each other.
-For instance, sequential palettes (except for spectral/rainbow palettes) all start with the lightest color and end with the darkest color.
-Furthermore, black and white are removed from most categorical palettes for practical reasons: white or light grey is almost always used as background color and black for annotation.
+*	Bivariate color palettes are available (besides the three main palette types: categorical, sequential, and diverging).
 
-* The GUI enables users to analyse characteristics of color palettes, such as color-blind-friendliness, harmony, and the presence of intense colors.
+*	Own color palettes can be loaded and analysed.
 
-* There is native support for **ggplot2** and **tmap** (as of the upcoming version 4).
+*	Color for missing values are made explicit.
 
-* It will be possible to submit an own series of color palettes.
+*   Palettes are made consistent with each other to enable comparison.
+For instance, black and white are (by default) removed from categorical palettes.
+Another standard that we adapt to is that all sequential palettes go from light to dark and not the other way round.
+
+*   There is native support for **ggplot2** and **tmap** (as of the
+    upcoming version 4).
+
+*	There are a couple of exporting options, including (bibtex) citation.
 

### 2023-01-05 R/c4a_name_convert.R

@@ -26,6 +26,16 @@ check_name_presence = function(name, verbose = TRUE) {
 	NULL
 }
 
+closest_match = function(x, y, tolerance = .5, max = 1.5) {
+	s = stringdist::stringdist(x, y, weight = c(d = .5, i = .7, s = 1, t = .7))
+	smin = min(s)
+	th = min(max, (smin + tolerance))
+	ids = which(s <= th)
+	list(ids = ids,
+		 stringdist = smin)
+}
+
+
 fuzzy_names = function(name) {
 	fullnames = .C4A$z$fullname
 	nnames = .C4A$z$name
@@ -36,19 +46,18 @@ fuzzy_names = function(name) {
 
 	# without "series."
 	if (length(base) == 0L) {
-		ids = which(stringdist::stringdist(name, nnames) <= 2)
-		message("\"", paste(fullnames[ids], collapse = "\", \""), "\"")
+		fullnames[closest_match(name, nnames)$ids]
 	} else {
-		ids_b = which(stringdist::stringdist(base, nnames[snames != series]) <= 1)
-		ids_ss = which(stringdist::stringdist(base, nnames[snames == series]) <= 2)
+		res1 = closest_match(base, nnames[snames != series])
+		res2 = closest_match(base, nnames[snames == series])
 
-		if (length(ids_b)) message("\"", paste(fullnames[snames == series][ids_ss], collapse = "\", \""), "\"")
-		if (length(ids_ss)) message("\"", paste(fullnames[snames != series][ids_b], collapse = "\", \""), "\"")
+		if (res2$stringdist <= res1$stringdist) {
+			fullnames[snames == series][res2$ids]
+		} else {
+			c(fullnames[snames == series][res2$ids],
+			  fullnames[snames != series][res1$ids])
+		}
 	}
-
-	which(stringdist::stringdist(name, nnames) <= 3)
-	stringdist::stringdist(name, fullnames)
-
 }
 
 #' Converts a name to a known palette name, if it exists.
@@ -70,9 +79,12 @@ c4a_name_convert = function(name, no.match = c("message", "error", "null"), verb
 		if (no.match == "error") {
 			stop("Unknown palette name. See c4a_palettes() for all palette names", call. = FALSE)
 		} else if (no.match == "message") {
-			message("Unknown palette name. Do you mean one of the following alternatives:")
-			fuzzy_names(name)
-			message("See c4a_palettes() for all palette names")
+			fuzzy = fuzzy_names(name)
[…]

### 2023-01-05 R/onLoad.R

@@ -1,22 +1,23 @@
 #' cols4all overview
 #'
-#' cols4all stands for: color palettes for all people, including those with color vision deficiency. Popular color palette series, such as ColorBrewer, have been organized by series and type and have been scored on several properties such as color-blind-friendliness and fairness. By default 436 palettes from 16 series are included, but own palettes can also be loaded and analysed. Besides the common palette types categorical, sequential, and diverging it also includes bivariate color palettes. ggplot2 scales are included.
+#' cols4all stands for: color palettes for all people, including those with color vision deficiency. Popular color palette series, such as ColorBrewer, have been organized by type and have been scored on several properties such as color-blind-friendliness and fairness (i.e. do colors stand out equally?). 436 color palettes from 16 series are included, but own palettes can also be loaded and analysed. Besides the common palette types (categorical, sequential, and diverging) it also includes bivariate color palettes. ggplot2 scales are included.
 #'
 #' This page provides a brief overview of all package functions.
 #'
 #' @section Main functions:
 #' \tabular{ll}{
-#' \code{\link{c4a_gui}}\tab Dashboard to analyse the palettes \cr
-#' \code{\link{c4a}}\tab Get the colors of a palette \cr
+#' \code{\link{c4a_gui}}\tab Dashboard for analyzing the palettes \cr
+#' \code{\link{c4a}}\tab Get the colors from a palette \cr
 #' }
 #'
 #' @section Palette names and properties:
 #' \tabular{ll}{
 #' \code{\link{c4a_palettes}}\tab Get available palette names \cr
 #' \code{\link{c4a_series}}\tab Get available series names\cr
-#' \code{\link{c4a_info}}\tab Get palette information from a specific palette (such as type and maximum number of colors) \cr
 #' \code{\link{c4a_citation}}\tab Show how to cites palettes (with bibtex code) \cr
+#' \code{\link{c4a_info}}\tab Get information from a palette, such as type and maximum number of colors) \cr
 #' \code{\link{.P}}\tab Environment via which palette names can be browsed with auto-completion (using `$`) \cr
+
 #' }
 #'
 #' @section Importing and exporting palettes:
@@ -193,9 +194,9 @@ c4a_options = function(...) {
 					inter_wing_dist = "Inter-wing-distance",
 					Crel = "Chroma (rel) max",
 					Cmax = "Chroma max",
-					H = "Hue mean",
[…]

### 2023-01-10 README.Rmd
fixed Blues-color selection bug

@@
 ```{r, eval=FALSE}
+# colorspace 2.1 required:
+install.packages(https://rdrr.io/r/utils/install.packages.html)("colorspace", repos = "https://R-Forge.R-project.org")
+
+# development (github) version of cols4all
 install.packages("remotes")

### 2023-01-10 R/c4a_gui.R

@@ -668,10 +668,10 @@ c4a_gui = function(type = "cat", n = NA, series = "all") {
 				tab_vals$colB1 = cols[1]
 				tab_vals$colB2 = cols[2]
 				tab_vals$CR = colorspace::contrast_ratio(cols[1], cols[2])
-				tab_vals$colC1 = cols[1]
-				tab_vals$colC2 = cols[2]
 				tab_vals$b = approx_blues(cols)
 				tab_vals$r = approx_reds(cols)
+				tab_vals$colC1 = cols[which.max(tab_vals$b)]
+				tab_vals$colC2 = cols[which.max(tab_vals$r)]
 
 			} else {
 				tab_vals$pal = character(0)
@@ -1170,7 +1170,6 @@ c4a_gui = function(type = "cat", n = NA, series = "all") {
 
 			ids = c(which.max(b)[1], which.max(r)[1])
 
-
 			#br = get_blue_red()
 
 			if (max(b) > .C4A$Blues && max(r) > 1) {
@@ -1203,6 +1202,7 @@ c4a_gui = function(type = "cat", n = NA, series = "all") {
 			if (!length(pal)) return(NULL)
 
 			cols = c(tab_vals$colC1, tab_vals$colC2)
+
 
 			lL = c(LETTERS,letters)[c(which(pal == cols[1]),
 					which(pal == cols[2]))]

### 2023-01-10 R/c4a_plot_X_matrix.R

@@ -124,7 +124,7 @@ plot_matrix = function(p, id1 = NULL, id2 = NULL, type = c("CR", "dist"), cvd = 
 
 				cellplot(i+1,j+1, {
 					grid::grid.points(x = 0.5, y = 0.5, pch = pchs[s], size = grid::unit(sizes[s], units = "lines"), gp = grid::gpar(col = fc))
-					if (!is.null(id1) && !is.null(id2) && id1 == i && id2 == j) {
+					if (!is.null(id1) && !is.null(id2) && id1[1] == i && id2[1] == j) {
 						grid::grid.circle(r = 0.4, gp = grid::gpar(fill = NA, col = fc, lwd = 1.5, lty = "dotted"))
 					}
 				})

### 2023-01-10 R/show_attach_scores.R

@@ -60,9 +60,9 @@ show_attach_scores = function(z) {
 					 ifelse(z2$HwidthL < .C4A$HwidthDivSingle & z2$HwidthR < .C4A$HwidthDivSingle, "SH", "MH"))
 		z2$HwidthLR = pmax(z2$HwidthL, z2$HwidthR)
 	}
-	z2$contrast = z2$CRmin <= 120
-	z2$contrastWT = z2$CRwt <= 120
-	z2$contrastBK = z2$CRbk <= 120
+	z2$contrast = z2$CRmin <= .C4A$contrastTH
+	z2$contrastWT = z2$CRwt <= .C4A$contrastTH
+	z2$contrastBK = z2$CRbk <= .C4A$contrastTH
 
 	z2$float = z2$Blues >= .C4A$Blues
 

### 2023-01-18 R/c4a_gui.R
cleaned up; CRAN check ok

@@ -1091,7 +1091,7 @@ c4a_gui = function(type = "cat", n = NA, series = "all") {
 			list(x = x_id, y = y_id)
 		}
 
-		observeEvent(input$cbfSimi_click, {
+		shiny::observeEvent(input$cbfSimi_click, {
 			pal = tab_vals$pal
 			if (!length(pal)) return(NULL)
 
@@ -1101,7 +1101,7 @@ c4a_gui = function(type = "cat", n = NA, series = "all") {
 			if (!is.na(ids$y)) tab_vals$colA1 = pal[ids$y]
 		})
 
-		observeEvent(input$cbfPSimi1_click, {
+		shiny::observeEvent(input$cbfPSimi1_click, {
 			pal = tab_vals$pal
 			if (!length(pal)) return(NULL)
 
@@ -1112,7 +1112,7 @@ c4a_gui = function(type = "cat", n = NA, series = "all") {
 
 		})
 
-		observeEvent(input$cbfPSimi2_click, {
+		shiny::observeEvent(input$cbfPSimi2_click, {
 			pal = tab_vals$pal
 			if (!length(pal)) return(NULL)
 
@@ -1122,7 +1122,7 @@ c4a_gui = function(type = "cat", n = NA, series = "all") {
 			if (!is.na(ids$y)) tab_vals$colA1 = pal[ids$y]
 		})
 
-		observeEvent(input$cbfPSimi3_click, {
+		shiny::observeEvent(input$cbfPSimi3_click, {
 			pal = tab_vals$pal
 			if (!length(pal)) return(NULL)
 
@@ -1133,7 +1133,7 @@ c4a_gui = function(type = "cat", n = NA, series = "all") {
 		})
 
 
-		observeEvent(input$table_click, {
+		shiny::observeEvent(input$table_click, {
 			pal = tab_vals$palBW
 			if (!length(pal)) return(NULL)
 
@@ -1142,7 +1142,7 @@ c4a_gui = function(type = "cat", n = NA, series = "all") {
 			if (!is.na(ids$x)) tab_vals$colB2 = pal[ids$x]
 			if (!is.na(ids$y)) tab_vals$colB1 = pal[ids$y]
 
-			if (!is.na(ids$x) || !is.na(id$y)) tab_vals$CR = colorspace::contrast_ratio(tab_vals$colB1, tab_vals$colB2)
[…]

### 2023-01-18 R/c4a_plot_hues.R

@@ -41,135 +41,135 @@ c4a_plot_hues = function(pal, dark = FALSE, C_grey = 5, width = c("none", "total
 
 
 	bc = if (dark) "#000000" else "#FFFFFF"
-	fc = if (dark) "#FFFFFF" else "#000000"
+		fc = if (dark) "#FFFFFF" else "#000000"
 
 
 
 
 
-	grid::grid.newpage()
-	grid::pushViewport(grid::viewport(width = grid::unit(1, "snpc"), height = grid::unit(1, "snpc")))
+		grid::grid.newpage()
+		grid::pushViewport(grid::viewport(width = grid::unit(1, "snpc"), height = grid::unit(1, "snpc")))
 
-	grid::grid.rect(gp = grid::gpar(fill = bc, col = NA))
+		grid::grid.rect(gp = grid::gpar(fill = bc, col = NA))
 
-	cx = 0.5
-	cy = 0.55
-	rd = 0.4
-	necklace_width = 0.02
+		cx = 0.5
+		cy = 0.55
+		rd = 0.4
+		necklace_width = 0.02
 
-	# plot necklace
+		# plot necklace
 
-	np = 5000
-	necklace = data.frame(h = runif(np, min = 0, max = 360),
-					 c = runif(np, min = .3, max = 1),
-					 l = runif(np, min = 30, max = 90))
+		np = 5000
+		necklace = data.frame(h = runif(np, min = 0, max = 360),
+							  c = runif(np, min = .3, max = 1),
+							  l = runif(np, min = 30, max = 90))
 
-	necklace$max_c = colorspace::max_chroma(h = necklace$h, l = necklace$l)
-	necklace$c = necklace$c * necklace$max_c
+		necklace$max_c = colorspace::max_chroma(h = necklace$h, l = necklace$l)
+		necklace$c = necklace$c * necklace$max_c
 
-	cols = hcl(necklace$h, necklace$c, necklace$l)
+		cols = hcl(necklace$h, necklace$c, necklace$l)
 
-	d = runif(np, min = rd - necklace_width/2, max = rd + necklace_width/2)
-	x = sin(necklace$h / 180 * pi) * d + cx
-	y = cos(necklace$h / 180 * pi) * d + cy
[…]

### 2023-01-18 R/c4a_table.R

@@ -54,10 +54,12 @@ table_columns = function(type, show.scores) {
 #' @param columns number of columns. By default equal to `n` or, if not specified, 12. Cannot be higher than the palette
 #' @param verbose should messages and warnings be printed?
 #' @import colorspace abind
-#' @importFrom grDevices hcl
+#' @importFrom grDevices hcl dev.size
 #' @importFrom stats lm predict
-#' @importFrom utils tail
+#' @importFrom utils tail head install.packages menu
 #' @importFrom png readPNG
+#' @importFrom stats rnorm runif
+#' @importFrom graphics barplot lines par
 #' @example ./examples/c4a_table.R
 #' @seealso References of the palettes: \code{\link{cols4all-package}}.
 #' @export
@@ -358,6 +360,3 @@ c4a_table = function(type = c("cat", "seq", "div", "bivs", "bivc", "bivd", "bivg
 	k
 
 }
-
-
-

### 2023-01-18 R/check_pals.R

@@ -362,7 +362,7 @@ get_CRmin = function(p, show.which = FALSE) {
 	if (show.which) {
 		message("id:", names(CRs)[id])
 		ids = as.integer(strsplit(names(CRs)[id], split = "_", fixed = TRUE)[[1]])
-		plus_rev(p[ids[1]], p[ids[2]])
+		c4a_plot_Plus_Reversed(p[ids[1]], p[ids[2]])
 	}
 	unname(CRs[id])
 	#structure(CRs[id], ids = strsplit(names(CRs[id]), split = "_", fixed = TRUE))

### 2023-01-18 R/onLoad.R

@@ -78,7 +78,7 @@ c4a_options = function(...) {
 		if (is.list(arg)) {
 			## case 1: option list is given
 			args = arg
-			if (length(lst) > 1 && show.warnings) warning("Only the first argument is used; the other arguments are ignored.")
+			if (length(lst) > 1) warning("Only the first argument is used; the other arguments are ignored.")
 		} else {
 			## case 2: option name is given
 			args = sapply(lst, "[", 1)

### 2023-01-18 sandbox/cbsnl.R

@@ -39,5 +39,5 @@ dat = readRDS("sandbox/cbsnl.rds")
 c4a_load(dat)
 
 \dontrun{
-c4a_gui(series = "cbsnl")
+c4a_gui(series = "cbsnl", n = 8)
 }

### 2023-01-24 README.Rmd
updated readme
...

@@
 # colorspace 2.1 required:
-install.packages(https://rdrr.io/r/utils/install.packages.html)("colorspace", repos = "https://R-Forge.R-project.org")
+install.packages("colorspace", repos = "https://R-Forge.R-project.org")
 
@@
 *   `c4a_series` Get available series names
+*   `c4a_types` Get implemented types
+* 	`c4a_overview` Get an overview of palettes per series x type.
 *   `c4a_citation` Show how to cite palettes (with bibtex code).

### 2023-01-24 R/onLoad.R

@@ -316,12 +316,12 @@ c4a_options = function(...) {
 			 contrastWT = list('NA' = "",
 			 				  'FALSE' =  "",
 			 				  'TRUE' = kableExtra::cell_spec("&#127987;",
-			 				  							   tooltip = "Low contrast with white background  (see tab 'Contrast')",
+			 				  							   tooltip = "Low contrast with white (see tab 'Contrast')",
 			 				  							   escape = FALSE, extra_css = "font-size: 130%; vertical-align: -0.1em; line-height: 0px; margin-right: -10px;")),
 			 contrastBK = list('NA' = "",
 			 				  'FALSE' =  "",
 			 				  'TRUE' = kableExtra::cell_spec("&#127988;",
-			 				  							   tooltip = "Low contrast with black background  (see tab 'Contrast')",
+			 				  							   tooltip = "Low contrast with black (see tab 'Contrast')",
 			 				  							   escape = FALSE, extra_css = "font-size: 130%; vertical-align: -0.1em; line-height: 0px;")),
 			 float = list('NA' = "",
 			 			 'FALSE' = "",

### 2023-01-25 R/c4a_data.R
added include.na to c4a_plot, added checks

@@ -51,7 +51,7 @@ c4a_data = function(x, xNA = NA, types = "cat", series = "x", nmin = NA, nmax = 
 	# check color list
 	if (!is.list(x)) stop("x is not a list")
 	nms = names(x)
-	x = lapply(x, validate_colors)
+	x = lapply(x, validate_colors, name = "x", from_list = TRUE)
 
 	# number of palettes
 	k = length(x)
@@ -69,7 +69,7 @@ c4a_data = function(x, xNA = NA, types = "cat", series = "x", nmin = NA, nmax = 
 	for (arg in args) assign(arg, rep(get(arg), length.out = k), envir = environment())
 
 	# validate na colors
-	if (any(!is.na(xNA))) xNA[!is.na(xNA)] = validate_colors(xNA[!is.na(xNA)])
+	if (any(!is.na(xNA))) xNA[!is.na(xNA)] = validate_colors(xNA[!is.na(xNA)], name = "xNA")
 
 	# check types
 	types_supported = unname(.C4A$types)

### 2023-01-25 R/c4a_gui.R

@@ -89,6 +89,7 @@ c4a_gui = function(type = "cat", n = NA, series = "all") {
 
 	stopifnot(length(types_available) > 0L)
 	if (!(type %in% types_available)) {
+		warning("type \"", type, "\" is not available/known")
 		type = types_available[1]
 	}
 
@@ -838,7 +839,7 @@ c4a_gui = function(type = "cat", n = NA, series = "all") {
 			if (!length(tab_vals$pal)) return(NULL)
 
 			pal = tab_vals$pal
-			c4a_plot_cvd(pal, dark = input$dark)
+			c4a_plot_cvd(pal, dark = input$dark, include.na = tab_vals$na)
 		})
 
 		output$cbfHL = shiny::renderPlot({

### 2023-01-25 R/c4a_plot.R

@@ -4,15 +4,17 @@
 #'
 #' @param palette Palette name (see \code{\link{c4a}}) or a color vector
 #' @param ... arguments passed on to \code{\link{c4a}}
+#' @param include.na should a color for missing values be included?
 #' @export
-c4a_plot = function(palette, ...) {
+c4a_plot = function(palette, ..., include.na = FALSE) {
 	args = list(...)
 
 	if (length(palette) == 1L) {
 		pal = do.call(c4a, c(list(palette = palette), args))
+		if (include.na) pal = c(pal, c4a_na(palette))
 	} else {
-		pal = palette
+		pal = validate_colors(palette, name = "palette")
 	}
 
-	c4a_plot_cvd(pal)
+	c4a_plot_cvd(as.vector(pal), include.na = include.na)
 }

### 2023-01-25 R/c4a_plot_cvd.R

@@ -1,4 +1,4 @@
-c4a_plot_cvd = function(cols, dark = FALSE, annotation = FALSE) {
+c4a_plot_cvd = function(cols, dark = FALSE, annotation = FALSE, include.na = FALSE) {
 	grid::grid.newpage()
 
 	fc = ifelse(dark, "#FFFFFF", "#000000")
@@ -17,7 +17,8 @@ c4a_plot_cvd = function(cols, dark = FALSE, annotation = FALSE) {
 
 	for (i in 2:(n+1)) {
 		cellplot(1, i, {
-			grid::grid.text(i-1, gp = grid::gpar(col = fc))
+			txt = if (i == (n+1) && include.na) "Missing" else (i - 1)
+			grid::grid.text(txt, gp = grid::gpar(col = fc))
 		})
 		for (j in 2:5) cellplot(j, i, {
 			grid::grid.rect(height = 0.8, gp=grid::gpar(fill = cols_lst[[j-1]][i-1], col = bc, lwd = 2))

### 2023-01-25 R/c4a_table.R

@@ -314,7 +314,12 @@ c4a_table = function(type = c("cat", "seq", "div", "bivs", "bivc", "bivd", "bivg
 	k = kableExtra::kbl(e2[, e2cols], col.names = e2th, escape = F)
 
 	for (cN in colNames) {
-		k = kableExtra::column_spec(k, which(cN == e2nms), width_min = "6em", width_max = "6em")
+		if (cN == " ") {
+			# column between colors and NA-color
+			k = kableExtra::column_spec(k, which(cN == e2nms), width_min = "1em", width_max = "1em")
+		} else {
+			k = kableExtra::column_spec(k, which(cN == e2nms), width_min = "6em", width_max = "6em")
+		}
 	}
 	for (i in which(substr(e2cols, 1, 4) == "Copy")) {
 		k = kableExtra::column_spec(k, i, width = "1em", extra_css = "padding-left: 10px; padding-right: 0px; text-align: right") #width_min = "1em", width_max = "1em")

### 2023-01-25 R/validate_colors.R

@@ -5,9 +5,16 @@ col2hex <- function(x) {
 }
 
 
-validate_colors <- function(x) {
-	if (!is.character(x)) stop("items of x found that are not a character vector", call. = FALSE)
-	if (any(is.na(x))) stop("items of x found that contain NAs", call. = FALSE)
+validate_colors <- function(x, name = "x", from_list = FALSE) {
+
+	if (from_list) {
+		pre = paste("not every item in ", name)
+	} else {
+		pre = name
+	}
+
+	if (!is.character(x)) stop(pre, " is not a character vector", call. = FALSE)
+	if (any(is.na(x))) stop(pre, "contain NAs", call. = FALSE)
 
 	w = which(x %in% colors())
 

### 2023-01-27 README.Rmd
...

@@
 "Color for all" refers to our mission that colors should be usable for not just people with normal color vision, but also for people with color vision deficiency.
-Currently, this package contains palettes from several popular and lesser known color palette series: [Color Brewer](https://colorbrewer2.org), [Viridis](http://bids.github.io/colormap/), [Kovesi](https://colorcet.com/), [Paul Tol](https://personal.sron.nl/~pault/), [Scico](https://www.fabiocrameri.ch/colourmaps/), [Carto](https://carto.com/carto-colors/), [Tableau](https://www.tableau.com/about/blog/2016/7/colors-upgrade-tableau-10-56782), [Wes Anderson](https://github.com/karthik/wesanderson), and [Seaborn](https://seaborn.pydata.org/tutorial/color_palettes.html).
+Currently, this package contains palettes from several popular and lesser known color palette series.
 Own palettes series can be added as well.
@@
 
-What palettes are available?
+What types and series are available?
 
-```{r, output.lines = 10}
-# All loaded palettes
-c4a_palettes()
+```{r}
+c4a_types()
+
+c4a_series()
 ```
+
+How many palettes per type x series?
+
+```{r}
+c4a_overview()
+```
+
+What palettes are available, e.g diverging from the hcl series?
 
@@
 ```{r}
-# Select the palette "kelly" with 7 colors
-c4a("kelly", 7)
-
 # select purple green palette from the hcl series:
@@
 
+Plot these colors:
+
+```{r, fig.height = 2}
+c4a_plot("hcl.purple_green", 11, include.na = TRUE)
+```
+
+## Using cols4all palettes in ggplot2
+
+```{r}
+library(ggplot2)
+data("diamonds")
+diam_exp = diamonds[diamonds$price >= 15000, ]
+
+# discrete categorical scale
+ggplot(diam_exp, aes(x = carat, y = price, color = color)) +
+	geom_point(size = 2) +
+	scale_color_discrete_c4a_cat("carto.safe") +
+	theme_light()
+
+# continuous diverging scale
+ggplot(diam_exp, aes(x = carat, y = depth, color = price)) +
+	geom_point(size = 2) +
+	scale_color_continuous_c4a_div("wes.zissou1", mid = mean(diam_exp$price)) +
+	theme_light()
+```
+
 ## Overview of functions
@@
 *   `c4a_gui` Dashboard for analyzing the palettes
-*   `c4a` Get the colors from a palette
+*   `c4a` Get the colors from a palette (`c4a_na` for the associated color for missing values)
+*	`c4a_plot` Plot a color palette
 

### 2023-01-27 R/c4a_data.R

@@ -1,16 +1,8 @@
 #' Build and load palette data
 #'
-#' Build palette data. Both `c4a_data` and `c4a_data_as_is` build data palette. The difference is that the former processes the palettes (see below) whereas the latter takes the palette data as is. Data can subsequently be loaded into cols4all via `c4a_load`.
+#' Build palette data. Both `c4a_data` and `c4a_data_as_is` build data palette. The difference is that the former may restructure the palette colors (see details) whereas the latter takes the palette colors as they are. Data can subsequently be loaded into cols4all via \code{\link{c4a_load}}.
 #'
-#' In cols4all, palettes are organized by series and by type. The **series** or 'family' specifies where the palettes belong to. For instance `"brewer"` stands for the color palettes from ColorBrewer. In cols4all we distinguish 7 **types**:
-#'
-#' **`"cat"`** categorical
-#' **`"seq"`** sequential
-#' **`"div"`** diverging
-#' **`"bivs`** bivariate (sequential x sequential)
-#' **`"bivc`** bivariate (sequential x categorical)
-#' **`"bivd`** bivariate (sequential x diverging)
-#' **`"bivg`** bivariate (sequential x desaturated (grayscale))
+#' In cols4all, palettes are organized by series and by type. The **series** or 'family' specifies where the palettes belong to. For instance `"brewer"` stands for the color palettes from ColorBrewer. Run \code{\link{c4a_series}} to get an overview of loaded series. The **type** specifies what kind of palette it is; see \code{\link{c4a_types}} for a description of the implemented ones.
 #'
 #' This function structures the palette data, such that it is consistent with the other palette data. This includes:
 #'
@@ -39,6 +31,7 @@
 #' @param space color space in which interpolated colors are determined. Options: `"rgb"` (RGB) and `"Lab"` (CIE Lab).
 #' @param range_matrix_args list of lists, one for each palette. Each such list specifies the range of sequential and diverging palettes, in case they are not indexed. See details.
 #' @param bib bibtex reference in the form of a `utils::bibentry` object.
+#' @param description description of the series. If `series` contains multiple series (rather than one value), please specify a vector of the same length as `series`. See \code{\link{c4a_series}} for the descriptions of the currently loaded series.
 #' @param ... passed on to `c4a_data`
 #' @example ./examples/c4a_data.R
 #' @rdname c4a_data

### 2023-01-27 R/c4a_gui.R

@@ -70,7 +70,8 @@ c4a_gui = function(type = "cat", n = NA, series = "all") {
 	tab_nmax = tapply(z$nmax, INDEX = list(z$series, factor(z$type, levels = tps)), FUN = max)
 	tab_k = as.data.frame(tapply(z$nmin, INDEX = list(z$series, factor(z$type, levels = tps)), FUN = length))
 	tab_k$series = rownames(tab_k)
-	tab_k = tab_k[, c("series", tps)]
+	tab_k$description = unname(.C4A$zdes[tab_k$series])
+	tab_k = tab_k[, c("series", "description", tps)]
 
 
 	allseries = sort(unique(z$series))
@@ -546,7 +547,8 @@ c4a_gui = function(type = "cat", n = NA, series = "all") {
 												shiny::renderTable(tab_k, na = "", striped = TRUE, hover = TRUE, bordered = TRUE),
 												shiny::div(style="font-size: 75%;", shiny::renderTable(.C4A$type_info)),
 												footer = shiny::modalButton("Close"),
-												style = "color: #000000;"))
+												style = "color: #000000;",
+												size = "l"))
 		})
 
 		get_cols = shiny::reactive({

### 2023-01-27 R/c4a_name_convert.R

@@ -60,13 +60,13 @@ fuzzy_names = function(name) {
 	}
 }
 
-#' Converts a name to a known palette name, if it exists.
-#'
-#' Converts a name to a known palette name, if it exists. Otherwise, either an error is thrown or `NULL` is returned
-#'
-#' @param name palette name to be converted
-#' @param no.match what happens is no match is found? Options: `"message"`: a message is thrown with suggestions, `"error"`: an error is thrown, `"null"`: `NULL` is returned
-#' @param verbose should messages be printed?
+# Converts a name to a known palette name, if it exists.
+#
+# Converts a name to a known palette name, if it exists. Otherwise, either an error is thrown or `NULL` is returned
+#
+# @param name palette name to be converted
+# @param no.match what happens is no match is found? Options: `"message"`: a message is thrown with suggestions, `"error"`: an error is thrown, `"null"`: `NULL` is returned
+# @param verbose should messages be printed?
 c4a_name_convert = function(name, no.match = c("message", "error", "null"), verbose = TRUE) {
 	no.match = match.arg(no.match)
 	name2 = check_name_presence(name, verbose)

### 2023-01-27 R/c4a_palettes.R

@@ -5,6 +5,7 @@
 #' @param type type of color palette: one of `"all"` (all palettes), `"cat"` (categorical/qualitative palettes), `"seq"` (sequential palettes) and `"div"` (diverging palettes).
 #' @param series series to list the palettes from. Run `c4a_series` to see the options.
 #' @param full.names should full names, i.e. with the prefix "series."? By default `TRUE`.
+#' @param as.data.frame should `c4a_series` and `c4a_types` return the result as a data.frame, with description included as a column?
 #' @return names of the loaded color palettes
 #' @example ./examples/c4a_palettes.R
 #' @seealso References of the palettes: \code{\link{cols4all-package}}.
@@ -36,7 +37,7 @@ c4a_series = function(type = c("all", "cat", "seq", "div"), as.data.frame = TRUE
 		return(invisible(NULL))
 	}
 	series = z$series
-	x = unique({if (type != "all") series[z$type == type] else series})
+	x = sort(unique({if (type != "all") series[z$type == type] else series}))
 
 	if (as.data.frame) {
 		data.frame(series = x, description = unname(.C4A$zdes[x]))

### 2023-01-27 R/c4a_sysdata_import.R

@@ -9,7 +9,7 @@
 #' @name c4a_sysdata_import
 #' @export
 c4a_sysdata_import = function(data) {
-	if (!is.list(data) || !setequal(c("data", "scores", "citation"), names(data))) stop("data should be a list of three: data, scores, citation", call. = FALSE)
+	if (!is.list(data) || !setequal(c("data", "scores", "citation", "description"), names(data))) stop("data should be a list of four: data, scores, citation, and description", call. = FALSE)
 
 	z = check_z(data$data)
 	s = check_s(data$scores, nrow(z))

### 2023-01-27 R/c4a_table.R

@@ -41,17 +41,17 @@ table_columns = function(type, show.scores) {
 #'
 #' Graphical user interface to analyse palettes. `c4a_table` shows a table that can be opened in the browser. `c4a_gui` is a graphical user interface (shiny app) around this table.
 #'
-#' @param type type of palette: `"cat"` for categorical (aka qualitative), `"seq"` for sequential, `"div"` for diverging, and `"bivs"`/`"bivc"`/`"bivd"`/`"bivg"` for bivariate (seq-seq, seq-cat, seq-div, and uncertainty-seq). For `c4a_gui` it only determines which type is shown initially.
+#' @param type type of palette. Run \code{\link{c4a_types}} to see the implemented types and their description. For `c4a_gui` it only determines which type is shown initially.
 #' @param n,m `n` is the number of displayed colors. For bivariate palettes `"biv"`, `n` and `m` are the number of columns and rows respectively. If omitted: for `"cat"` the full palette is displayed, for `"seq"` and `"div"`, 9 colors, and for `"bivs"`/`"bivc"`/`"bivd"`/`"bivg"` 4 columns and rows. For `c4a_gui` it only determines which number of colors initially.
 #' @param cvd.sim color vision deficiency simulation: one of `"none"`, `"deutan"`, `"protan"`, `"tritan"`
-#' @param sort column name to sort the data. For column names, see details. Use a `"-"` prefix to reverse the order.
+#' @param sort column name to sort the data. The available column names depend on the arguments `type` and `show.scores`. They are listed in the warning message. Use a `"-"` prefix to reverse the order.
 #' @param text.format The format of the text of the colors. One of `"hex"`, `"RGB"` or `"HCL"`.
 #' @param text.col The text color of the colors. By default `"same"`, which means that they are the same as the colors themselves (so invisible, but available for selection). `"auto"` means automatic: black for light colors and white for dark colors.
 #' @param series Series of palettes to show. See \code{\link{c4a_series}} for options. By default, `"all"`, which means all series. For `c4a_gui` it only determines which series are shown initially.
[…]

### 2023-01-27 examples/c4a.R

@@ -3,7 +3,8 @@ c4a_palettes("div")
 c4a(type = "cat")
 
 (pal = c4a("tol.sunset", n = 7, range = c(0, .6)))
-colorspace::specplot(pal)
+
+c4a_plot(pal)
 
 c4a("set2")
 
@@ -16,3 +17,6 @@ c4a("hcl.set2", reverse = TRUE, n = 8)
 
 # handy shortcut
 c4a("-hcl.set2", n = 8)
+
+# the color for missing values is white:
+c4a_na("hcl.set2")

### 2023-01-27 examples/c4a_data.R

@@ -1,22 +1,39 @@
-## palettes from Pink Floyd albums
-pf = list(piper = c("#391C1C", "#C6C6AA", "#713939", "#C6391C", "#C6E3C6", "#AA7155", "#AA8E71", "#C68E71"),
-		  saucerful = c("#000000", "#1C1C1C", "#393939", "#FFFFFF", "#555555", "#8E8E71", "#E3C6AA", "#715539"),
-		  atom = c("#C6E3FF", "#397139", "#557139", "#E3E3C6", "#1C1C1C", "#1C551C", "#AAAA8E", "#8EC6E3"),
-		  meddle = c("#715539", "#553939", "#8E7155", "#71AAAA", "#8E8E71", "#1CAAE3", "#55C6E3", "#AA7155"),
-		  obscured = c("#000000", "#1C1C1C", "#393939", "#717155", "#8E8E71", "#715539", "#C6AA8E", "#E3C6AA"),
-		  moon = c("#000000", "#FF0000", "#FF9224", "#FFFF00", "#71C600", "#00C6FF", "#8E398E", "#FFFFFF"),
-		  wish = c("#FFFFFF", "#AAC6E3", "#8E8E8E", "#717155", "#555539", "#8E8E71", "#555555", "#8E7155"),
-		  animals = c("#391C39", "#393955", "#E3C671", "#718E8E", "#AAAA8E", "#C67139", "#AA5539", "#E3AA39"),
-		  wall = c("#FFFFFF", "#E3E3E3", "#C6C6C6", "#AAAAC6", "#1C1C1C", "#000000", "#8E8E8E", "#E3C6E3"),
-		  cut = c("#000000", "#E30000", "#AA0000", "#391C55", "#FFE3E3", "#1C1C00", "#FFAA55", "#8E8E55"),
-		  lapse = c("#000000", "#8E8EC6", "#8E8E71", "#7171AA", "#39391C", "#717171", "#AAAAAA", "#E3E3E3"),
-		  division = c("#000000", "#FFFFC6", "#00398E", "#AA8E55", "#39558E", "#C6AA71", "#39391C", "#555571"),
-		  more = c("#0055AA", "#FFAA1C", "#1C71AA", "#003971", "#E38E55", "#E3AAAA", "#718EAA", "#71718E"),
-		  umma = c("#AA8E71", "#555539", "#39391C", "#1C1C1C", "#E3E3C6", "#715539", "#391C1C", "#8E7155"),
[…]

### 2023-01-29 R/c4a_citation.R
vignette doesn't appear yet

@@ -5,9 +5,10 @@
 #' @param name name of a palette or series
 #' @param verbose should text be printed (if FALSE only a `utils::bibentry` object is returned)
 #' @return `utils::bibentry` object
+#' @example ./examples/c4a_citation.R
 #' @export
 c4a_citation = function(name, verbose = TRUE) {
-	is_series = name %in% c4a_series()
+	is_series = name %in% c4a_series(as.data.frame = FALSE)
 
 	if (!is_series) {
 		inf = c4a_info(name, no.match = NULL, verbose = FALSE)

### 2023-01-29 examples/c4a_citation.R

@@ -1,0 +1,3 @@
+c4a_citation("hcl")
+
+c4a_citation("poly.glasbey")

### 2023-02-02 README.Rmd
CRAN 0.6

@@
 
+[![CRAN](http://www.r-pkg.org/badges/version/cols4all)](https://cran.r-project.org/package=cols4all) 
+[![cran checks](https://cranchecks.info/badges/worst/cols4all)](https://cran.r-project.org/web/checks/check_results_cols4all.html)
+[![Downloads](http://cranlogs.r-pkg.org/badges/cols4all?color=brightgreen)](http://www.r-pkg.org/pkg/cols4all)
+
 The **cols4all** is a new R package for selecting color palettes.
@@
 
-**cols4all** will be available on CRAN in January or February this year (2023). Until then it can be installed using:
+**cols4all** is available on CRAN:
 
 ```{r, eval=FALSE}
-# colorspace 2.1 required:
-install.packages("colorspace", repos = "https://R-Forge.R-project.org")
+install.packages("cols4all", dependencies = TRUE)
+```
 
-# development (github) version of cols4all
+The development version can be installed as follows:
+
+```{r, eval=FALSE}
 install.packages("remotes")

### 2023-02-02 README.Rmd
added paletteer demo

@@
 Let us know! (via github issues)
+
+## Bonus: 2568 palettes from **paletteer**
+
+The R package [**paletteer**](https://github.com/EmilHvitfeldt/paletteer) contains 2568 from 70 other R packages.
+These have been converted to data for **cols4all** [(script)](https://github.com/mtennekes/cols4all/blob/main/build/build_paletteer.R)
+
+These can be loaded and analysed as follows:
+
+```{r}
+paletteer_data = readRDS(url("https://github.com/mtennekes/cols4all/releases/download/v0.6/paletteer.rds","rb"))
+c4a_sysdata_import(paletteer_data)
+```
+
+Note that we use `c4a_sysdata_import` rather than `c4a_load` because we want to replace all the palette data to avoid overlap and conflicts with the palettes already included in **cols4all**.
+
+```{r}
+c4a_overview()
+```
+
+```{r, eval=FALSE}
+c4a_gui()
+```
+
+![https://user-images.githubusercontent.com/2444081/216455415-de676d12-8e94-4917-8adc-8ca1e6ac4df7.png](https://user-images.githubusercontent.com/2444081/216455415-de676d12-8e94-4917-8adc-8ca1e6ac4df7.png)
+
+Note that this dashboard is not particularly fast (so grab a coffee:-))

### 2024-03-02 README.Rmd
updated vignette, struggling with figure

@@
 
-The **cols4all** is a new R package for selecting color palettes.
+**cols4all** is an R package for selecting color palettes.
 "Color for all" refers to our mission that colors should be usable for not just people with normal color vision, but also for people with color vision deficiency.
@@
 
-![https://user-images.githubusercontent.com/2444081/210850914-cdb8a128-1b8a-4900-94d0-dfa6ca449585.png](https://user-images.githubusercontent.com/2444081/210850914-cdb8a128-1b8a-4900-94d0-dfa6ca449585.png)
+[<img src="figures/cols4all_table3.png" width="650"/>](figures/cols4all_table3.png)
 
@@
 
+Edit color palette data
+
+*   `c4a_duplicate` Duplicates a color palette
+*   `c4a_modify` Modifies palette colors
+
 ggplot2

### 2024-03-02 R/c4a_gui.R

@@ -333,7 +333,7 @@ c4a_gui = function(type = "cat", n = NA, series = "all") {
 							shiny::fluidRow(
 								shiny::column(width = 12,
 									shiny::selectizeInput("namePal", "Palette", choices = init_pal_list),
-									infoBoxUI("infoName", "Naming table"),
+									infoBoxUI("infoName", "Naming table (in development)"),
 									plotOverlay("anaName", width = "1000px", height = "600px", "aniName"))),
 									#shiny::plotOutput("namePlot", height = "600px", width = "1000px"))),
 							shiny::fluidRow(

### 2024-03-02 R/c4a_modify.R

@@ -1,9 +1,9 @@
-#' Edit cols4all palettes
+#' Edit cols4all palettes (in development)
 #'
 #' Edit cols4all palettes. c4a_duplicate duplicates an existing cols4all palette, and c4a_modify is used to change the colors. Use c4a_data to craete palettes from scratch.
 #'
 #' @param palette name of the palette
-#' @param x vector of the new colors. It should either the same length, or a named vector, where the names correspond to the index numbers. E.g. c("3" = "#AABBCC") will replace the third color with the color "#AABBCC".
+#' @param x vector of the new colors. It should either the same length, or a named vector, where the names correspond to the index numbers. E.g. `c("3" = "#AABBCC")` will replace the third color with the color `"#AABBCC"`.
 #' @param xNA the new color for missing values.
 #' @rdname c4a_modify
 #' @name c4a_modify

### 2024-03-04 README.Rmd

@@
 
-# Colors for all! <img src='inst/img/cols4all_logo.png' align="right" height="139" />
+# Colors for all! <a href="https://mtennekes.github.io/cols4all/"><img src="man/figures/logo.png" align="right" height="138" alt="cols4all website" /></a>
 
-[![CRAN](http://www.r-pkg.org/badges/version/cols4all)](https://cran.r-project.org/package=cols4all) 
+<!-- badges: start --->
+[![CRAN](https://www.r-pkg.org/badges/version/cols4all)](https://cran.r-project.org/package=cols4all) 
 [![cran checks](https://cranchecks.info/badges/worst/cols4all)](https://cran.r-project.org/web/checks/check_results_cols4all.html)
-[![Downloads](http://cranlogs.r-pkg.org/badges/cols4all?color=brightgreen)](http://www.r-pkg.org/pkg/cols4all)
+[![Downloads](https://cranlogs.r-pkg.org/badges/cols4all?color=brightgreen)](https://www.r-pkg.org/pkg/cols4all)
+<!-- badges: end --->
 

### 2024-03-06 README.Rmd
finetuning, testing images

@@
 
-[<img src="figures/cols4all_table3.png" width="650"/>](figures/cols4all_table3.png)
+[<img src="vignettes/dash-3.png" width="650"/>](vignettes/dash-3.png)
 
@@
 Let us know! (via github issues)
-
-## Bonus: 2568 palettes from **paletteer**
-
-The R package [**paletteer**](https://github.com/EmilHvitfeldt/paletteer) contains 2568 from 70 other R packages.
-These have been converted to data for **cols4all** [(script)](https://github.com/mtennekes/cols4all/blob/main/build/build_paletteer.R)
-
-These can be loaded and analysed as follows:
-
-```{r}
-paletteer_data = readRDS(url("https://github.com/mtennekes/cols4all/releases/download/v0.6/paletteer.rds","rb"))
-c4a_sysdata_import(paletteer_data)
-```
-
-Note that we use `c4a_sysdata_import` rather than `c4a_load` because we want to replace all the palette data to avoid overlap and conflicts with the palettes already included in **cols4all**.
-
-```{r}
-c4a_overview()
-```
-
-```{r, eval=FALSE}
-c4a_gui()
-```
-
-![https://user-images.githubusercontent.com/2444081/216455415-de676d12-8e94-4917-8adc-8ca1e6ac4df7.png](https://user-images.githubusercontent.com/2444081/216455415-de676d12-8e94-4917-8adc-8ca1e6ac4df7.png)
-
-Note that this dashboard is not particularly fast (so grab a coffee:-))

### 2024-03-06 R/c4a_gui.R

@@ -56,7 +56,6 @@ c4a_gui = function(type = "cat", n = NA, series = "all") {
 
 
 	shiny::addResourcePath(prefix = "imgResources", directoryPath = system.file("man/figures", package = "cols4all"))
-	shiny::addResourcePath(prefix = "imgResources", directoryPath = system.file("img", package = "cols4all"))
 
 	#############################
 	## Catelogue tab

### 2024-03-06 R/onLoad.R

@@ -274,8 +274,8 @@ do_cellspec = function(lst) {
 
 		tc = list(cbfriendly = list('NA' = "",
 									'0' = "",
-									'2' = list("&#9786;&#9786;", extra_css="font-size: 80%;", tooltip = "Colorblind-friendly!", escape = FALSE),
-									'1' = list("&#9786;", extra_css="font-size: 80%;", tooltip = "Colorblind-friendly!", escape = FALSE),
+									'2' = list("&#9786;&#9786;", extra_css="font-size: 80%;", tooltip = "Extra colorblind-friendly! Also for points and lines", escape = FALSE),
+									'1' = list("&#9786;", extra_css="font-size: 80%;", tooltip = "Colorblind-friendly! Be careful with points and lines", escape = FALSE),
 									'-1' = list("&#128064;", extra_css ="font-size: 60%;", tooltip = "Be careful! Some colors are hard to distinguish by color blind people (see tab 'Color Blind Friendliness'", escape = FALSE)),
 				  chroma = list('NA' = "",
 				  			  'H' = list("&#x1f576;", tooltip = "Vivid colors (high chroma): ideal for small important objects to stand out (e.g. markers on a map), but less suited for space filling visualizations (see tab 'HCL Analysis')", escape = FALSE),

### 2024-10-07 README.Rmd
updated readme

@@
 Finally, for each color palette a color for missing values is assigned, which is especially important for spatial data visualization.
-Currently we support several types: *categorical* (qualitative) palettes, *sequential* palettes, *diverging* palettes, and *bivariate* palettes (divided into four subtypes).
+Currently we support several types: *categorical* (qualitative) palettes, *sequential* palettes, *diverging* palettes, *cycling* palettes and *bivariate* palettes (divided into four subtypes).
 
@@
 ```
+
+This development version will be on CRAN soon (October 2024)
 
@@
 
-[<img src="vignettes/dash-3.png" width="650"/>](vignettes/dash-3.png)
+[<img src="vignettes/cols4all2.png" width="900"/>](vignettes/cols4all2.png)
 
-What types and series are available?
+What palettes are available? That is, by default; other palettes can be added!
 
 ```{r}
-c4a_types()
-
 c4a_series()
@@
 
-How many palettes per type x series?
+## Using the tool
 
-```{r}
-c4a_overview()
-```
+Use the tool to compare palettes and if needed analyse a palette in depth (via the other tabs).
 
-What palettes are available, e.g diverging from the hcl series?
+Find a trade-off you like among the following properties (the columns in the main table):
 
-```{r}
-# Diverging palettes from the 'hcl' series
-c4a_palettes(type = "div", series = "hcl")
-```
+* Colorblind friendly: Is the palette color blind friendly?
+* Fair: Is the palette fair? In a fair palette, colors stand out about equally. 
+* Hues: What hue ranges are used? All across the (rainbow) color spectrum, or a limited range?
+* Vivid: Are there any colors that are very saturated? Perhaps a little too much?
+* Contrast: Is there sufficient contrast against white, black (using WACG criteria) and between colors?
+* 3D Blues: If blue is a palette color, a 3D visual illusion could appear.
+* Naming (in development): How well can palette colors be named?
 
-Give me the colors!
+### Example 1
 
-```{r}
-# select purple green palette from the hcl series:
-c4a("hcl.purple_green", 11)
+When we are looking for a fair categorical palette of seven colors that is as color blind friendly as possible, then filter on "Fair", and sort by "Colorblind-friendly":
 
-# get the associated color for missing values
-c4a_na("hcl.purple_green")
-```
+[<img src="vignettes/cols4all_fair_sort_cbf.png" width="900"/>](vignettes/cols4all_fair_sort_cbf.png)
 
-Plot these colors:
+This gave us inspiration to develop own palettes: see these `cols4all` palettes below.
 
-```{r, fig.height = 2}
-c4a_plot("hcl.purple_green", 11, include.na = TRUE)
-```
+### Example 2
 
-## Using cols4all palettes in ggplot2
+Say we need a diverging palette that is color blind friendly, and what to choose one by eye. Then filter by "Colorblind-friendly" and sort by "Huddle Middle L" (the hue of the left wing):
+
+[<img src="vignettes/cols4all_div.png" width="900"/>](vignettes/cols4all_div.png)
+
+Reverse sorting is also applied.
+
+## Preliminary set of new `cols4all` palettes
+
+We applied a basic heuristics to explore palettes that score well on a mix of the properties named above
+
+[<img src="vignettes/c4a_pals.png" width="900"/>](vignettes/c4a_pals.png)
+
+`area7`, `area8` and `area9` are fair, contain low pastel colors, and are color-blind friendly (up to 7 colors).
+So ideal for maps and other space-filling visualizations!
+These are used in [https://github.com/r-tmap/tmap](`tmap4`).
+
+[<img src="vignettes/area7.png" width="900"/>](vignettes/area7.png)
+
+`area7d`, `area8d` and `area9d` similar but for dark mode:.
+
+[<img src="vignettes/area7d.png" width="900"/>](vignettes/area7d.png)
+
+`line7`, `line8` and `line9` are colors with good contrast against both black and white, and are also colorblind-friendly to some extent. So ideal for line graphs and scatter plots:
+
+[<img src="vignettes/line7_wt.png" width="440"/>](vignettes/line7_wt.png)
+[<img src="vignettes/line7_bk.png" width="440"/>](vignettes/line7_bk.png)
+
+Finally `friendly7` ... `friendly13` are colorblind-friendly palettes (disregarding the other properties):
+
+[<img src="vignettes/friendly13.png" width="900"/>](vignettes/friendly13.png)
+
+## `ggplot2` integration
 
@@
 
+## Other R funtions
+
+What palettes are available, e.g diverging from the hcl series?
+
+```{r}
+# Diverging palettes from the 'hcl' series
+c4a_palettes(type = "div", series = "hcl")
+```
+
+Give me the colors!
+
+```{r}
+# select purple green palette from the hcl series:
+c4a("hcl.purple_green", 11)
+
+# get the associated color for missing values
+c4a_na("hcl.purple_green")
+```
+
+Plot these colors:
+
+```{r, fig.height = 2}
+c4a_plot_cvd("hcl.purple_green", 11, include.na = TRUE)
+```
+
 ## Related R packages
@@
 
-*	Bivariate color palettes are available (besides the three main palette types: categorical, sequential, and diverging).
+*	Bivariate color palettes are available.
 
@@
 
-*	Color for missing values are made explicit.
-
-*   Palettes are made consistent with each other to enable comparison.
-For instance, black and white are (by default) removed from categorical palettes.
-Another standard that we adapt to is that all sequential palettes go from light to dark and not the other way round.
+*	Colors for missing values are made explicit.
 

### 2024-10-08 R/c4a.R
updated vignette and docs, updated carto (12th color as colNA)

@@ -3,14 +3,14 @@
 #' Get a cols4all color palette: `c4a` returns the colors of the specified palette, and `c4a_na` returns the color for missing value that is associated with the specified palette. Run \code{\link{c4a_gui}} to see all available palettes, which are also listed with \code{\link{c4a_palettes}}.
 #'
 #' @param palette name of the palette. See \code{\link{c4a_palettes}} for options. If omitted, the default palette is provided by `c4a_default_palette`. The palette name can be prefixed with a `"-"` symbol, which will reverse the palette (this can also be done with the `reverse` argument).
-#' @param n number of colors. If omitted then: for type `"cat"` the maximum number of colors is returned, for types `"seq"` and `"div"`, 9 colors.
+#' @param n number of colors. If omitted then: for type `"cat"` the maximum number of colors is returned, for types `"seq"`, `"div"`, and `"cyc"`, 7 , 9, and 9 colors respectively.
 #' @param m number of rows in case type is `"bivs"`, `"bivc"`, `"bivd"` or  `"bivg"` (which stand for respectively sequential, categorical, diverging and desaturated (`g` for 'grayscale')).
-#' @param type type of color palette, in case `palette` is not specified: one of `"cat"` (categorical/qualitative palette), `"seq"` (sequential palette), `"div"` (diverging palette), and `"bivs"`/`"bivc"`/`"bivd"`/`"bivg"` (bivariate: respectively seq-seq seq-cat, seq-div, and seq-desaturated).
+#' @param type type of color palette, in case `palette` is not specified: one of `"cat"` (categorical/qualitative palette), `"seq"` (sequential palette), `"div"` (diverging palette), `"cyc"` (cyclic palette), and `"bivs"`/`"bivc"`/`"bivd"`/`"bivg"` (bivariate: respectively seq-seq seq-cat, seq-div, and seq-desaturated).
 #' @param reverse should the palette be reversed?
 #' @param order order of colors. Only applicable for `"cat"` palettes
[…]

### 2024-10-08 R/c4a_palettes.R

@@ -1,8 +1,8 @@
 #' Get available palette names and series
 #'
-#' `c4a_palettes` lists all available cols4all color palettes. Palettes are organized by series. The available series are listed with `c4a_series`. Palettes are also organized per functional type, where we currently support: categorical `"cat"`, sequential `"seq"`, and diverging `"div"`" palette types. The function `c4a_types` lists all available types. The function `c4a_overview` gives an overview table of the number of palette per series and type. In an IDE with auto-completion (such as RStudio) it is possible to browse through the palette names with `.P` (using `$` like in lists).
+#' `c4a_palettes` lists all available cols4all color palettes. Palettes are organized by series. The available series are listed with `c4a_series`. Palettes are also organized per functional type, where we currently support: categorical `"cat"`, sequential `"seq"`, diverging `"div"`", cyclic `"cyc"`, and bivariate (seq x seq `"bivs"`, seq x cat `"bivc"`, seq x div `"bivd"`, seq x desaturated `"bivg"`) palette types. The function `c4a_types` lists all available types. The function `c4a_overview` gives an overview table of the number of palette per series and type. In an IDE with auto-completion (such as RStudio) it is possible to browse through the palette names with `.P` (using `$` like in lists).
 #'
-#' @param type type of color palette: one of `"all"` (all palettes), `"cat"` (categorical/qualitative palettes), `"seq"` (sequential palettes) and `"div"` (diverging palettes).
+#' @param type type of color palette: one of `"all"` (all palettes), `"cat"` (categorical/qualitative palettes), `"seq"` (sequential palettes), `"div"` (diverging palettes), `"cyc"` (cyclic), and one of the bivariate palette types: `"bivs"`, `"bivc"`, `"bivd"`, `"bivg"`
 #' @param series series to list the palettes from. Run `c4a_series` to see the options.
 #' @param full.names should full names, i.e. with the prefix "series."? By default `TRUE`.
 #' @param as.data.frame should `c4a_series` and `c4a_types` return the result as a data.frame, with description included as a column?

### 2024-10-08 R/c4a_sysdata_import.R

@@ -12,7 +12,16 @@ c4a_sysdata_import = function(data) {
 	if (!is.list(data) || !setequal(c("data", "scores", "citation", "description"), names(data))) stop("data should be a list of four: data, scores, citation, and description", call. = FALSE)
 
 	z = check_z(data$data)
-	s = check_s(data$scores, nrow(z))
+
+	svalid = check_s(data$scores, nrow(z))
+
+	s = if (svalid) {
+		data$scores
+	} else {
+		message("Recalculating scores")
+		series_add_get_scores(z)
+	}
+
 	zbib = data$citation # to do: check
 	zdes = data$description
 

### 2024-10-08 R/c4a_table.R

@@ -484,7 +484,7 @@ plot_table = function(p, text.format, text.col, include.na, cvd.sim, verbose) {
 #' Graphical user interface to analyse palettes. `c4a_table` shows a table that can be opened in the browser. `c4a_gui` is a graphical user interface (shiny app) around this table.
 #'
 #' @param type type of palette. Run \code{\link{c4a_types}} to see the implemented types and their description. For `c4a_gui` it only determines which type is shown initially.
-#' @param n,m `n` is the number of displayed colors. For bivariate palettes `"biv"`, `n` and `m` are the number of columns and rows respectively. If omitted: for `"cat"` the full palette is displayed, for `"seq"` and `"div"`, 9 colors, and for `"bivs"`/`"bivc"`/`"bivd"`/`"bivg"` 4 columns and rows. For `c4a_gui` it only determines which number of colors initially.
+#' @param n,m `n` is the number of displayed colors. For bivariate palettes `"biv"`, `n` and `m` are the number of columns and rows respectively. If omitted: for `"cat"` the full palette is displayed, for `"seq"`, `"div"` and `"cyc"`, 7, 9, and 9 colors respectively, and for `"bivs"`/`"bivc"`/`"bivd"`/`"bivg"` 4 columns and rows. For `c4a_gui` it only determines which number of colors initially.
 #' @param filters filters to be applied. A character vector with a subset from:`"nmax"` (only palettes where `n = nmax`, which is only applicable for categorical palettes),  `"cbf"` (colorblind-friendly), `"fair"` (fairness),`"naming"` (nameability),  `"crW"` (sufficient contrast ratio with white), and `"crB"` (sufficient contrast ratio with black). By default an empty vector, so no filters are applied.
 #' @param cvd.sim color vision deficiency simulation: one of `"none"`, `"deutan"`, `"protan"`, `"tritan"`
 #' @param sort column name to sort the data. The available column names depend on the arguments `type` and `show.scores`. They are listed in the warning message. Use a `"-"` prefix to reverse the order.

### 2024-10-08 R/onLoad.R

@@ -129,7 +129,7 @@ do_cellspec = function(lst) {
 	attach_bib()
 
 	with(.C4A,{
-		defaults = c(cat = "tol.muted", seq = "hcl.blues2", div = "hcl.purple_green", bivs = "c4a.bu_br_bivs", bivc = "met_monet", bivd = "c4a.pu_gn_bivd", bivg = "c4a.br_bivg")
+		defaults = c(cat = "cols4all.line7", seq = "kovesi.blue", div = "cols4all.pu_gn_div", cyc = "scico.roma_o", bivs = "cols4all.bu_br_bivs", bivc = "met_monet", bivd = "cols4all.pu_gn_bivd", bivg = "cols4all.br_bivg")
 
 		score_x100 = c("min_dist", "min_step", "max_step", "inter_wing_dist", "tri_ineq", "CRmin", "CRwt", "CRbk", "Blues")
 

### 2024-10-08 build/build_paletteer.R

@@ -1,12 +1,30 @@
 #reprex::reprex({
 sessionInfo(package = "paletteer")
+
+## VERSION 1.6
 library(paletteer)
 
+# continuous
 table(palettes_c_names$type)
-# typo: sequantial
 
-table(palettes_dynamic_names$type)
+
+palettes_c
+
+# discrete
 table(palettes_d_names$type)
+
+# dynamic fixed (i.e. indexed)
+table(palettes_dynamic_names$type)
+
+
+# discrete (fixed)			discrete (indexed)		continuous
+# palettes_d_names			palettes_dynamic_names  palettes_c_names
+
+
+
+######## d
+
+
 # divergent instead of diverging
 
 palettes_d$ggthemes$stata_economist
@@ -14,32 +32,30 @@ palettes_d$ggthemes$stata_economist
 
 # palettes_d is not consistent with palettes_d_names:
 pals_d = do.call(c, palettes_d)
-setdiff(names(pals_d), paste0(palettes_d_names$package, ".", palettes_d_names$palette))
-setdiff(paste0(palettes_d_names$package, ".", palettes_d_names$palette), names(pals_d))
-#})
-
-
-# discrete (fw)		discrete (dyn)			continuous
-# palettes_d_names	palettes_dynamic_names  palettes_c_names
-
 df_d = palettes_d_names
+
+setdiff(names(pals_d), paste0(df_d$package, ".", df_d$palette))
+setdiff(paste0(df_d$package, ".", df_d$palette), names(pals_d))
+
+# manual fix
+df_d$palette[df_d$palette == "hamadryas_feronia"] = c("hamadryas_feronia", "hamadryas_feronia_2")
+
+
+
+
+
 df_d$type = ifelse(df_d$type %in% c("sequantial", "sequential"), "seq", ifelse(df_d$type == "divergent", "div", "cat"))
 df_d$n = df_d$length
 
 setdiff(names(pals_d), paste0(df_d$package, ".", df_d$palette))
 setdiff(paste0(df_d$package, ".", df_d$palette), names(pals_d))
 
+pals_d2 = pals_d[match(paste0(df_d$package, ".", df_d$palette), names(pals_d))]
 
-df_d = df_d[df_d$palette != "contrast", ]
[…]

### 2024-10-08 build/build_palettes.R

@@ -527,15 +527,11 @@ local({
 local({
 	cartoQual = cartocolors[cartocolors$Type == "qualitative",]
 
-	indices = lapply(1:nrow(cartoQual), function(i) {
-		structure(lapply(1:12, function(j) {
-			if (j < 3) 1:j else match(cartoQual[[paste0("n", j)]][[i]], cartoQual$n12[[i]])
-		}), names = as.character(1:12))
-	})
+	# seems indexed, but only the 12th color (a gray) is added.
+	# -> use that 12th as colNA
 
-	pals = mapply(function(pal, ind) {
-		structure(pal, index = ind)
-	}, cartoQual$n12, indices, SIMPLIFY = FALSE)
+	pals = lapply(cartoQual$n12, function(p) p[1:11])
+	colNAs = sapply(cartoQual$n12, function(p) p[12])
 
 	names(pals) = cartoQual$Name
 
@@ -552,7 +548,7 @@ local({
 	names(pals3) = tolower(cartoAgg$Name)
 	names(pals3)[2] = "ag_grn_yl"
 
-	c4a_load(c4a_data(pals, types = "cat", series = "carto"))
+	c4a_load(c4a_data(pals, xNA = colNAs, types = "cat", series = "carto"))
 	c4a_load(c4a_data(pals2, types = type, series = "carto"))
 	c4a_load(c4a_data(pals3, types = "seq", series = "carto", format.palette.name = FALSE))
 

### 2024-10-08 build/sysdata_from_older_versions.R

@@ -1,0 +1,5 @@
+c71 = readRDS("sandbox/cols4all_71.rds")
+
+c4a_sysdata_import(c71)
+
+c71_8 = c4a_sysdata_export()

### 2024-10-09 R/c4a.R
0.8 ready

@@ -1,15 +1,15 @@
 #' Get a cols4all color palette
 #'
-#' Get a cols4all color palette: `c4a` returns the colors of the specified palette, and `c4a_na` returns the color for missing value that is associated with the specified palette. Run \code{\link{c4a_gui}} to see all available palettes, which are also listed with \code{\link{c4a_palettes}}.
+#' Get a cols4all color palette: `c4a` returns the colors of the specified palette, `c4a_na` returns the color for missing value that is associated with the specified palette, and `c4a_ramp` returns a color ramp function.  Run \code{\link{c4a_gui}} to see all available palettes, which are also listed with \code{\link{c4a_palettes}}.
 #'
-#' @param palette name of the palette. See \code{\link{c4a_palettes}} for options. If omitted, the default palette is provided by `c4a_default_palette`. The palette name can be prefixed with a `"-"` symbol, which will reverse the palette (this can also be done with the `reverse` argument).
+#' @param palette name of the palette. See \code{\link{c4a_palettes}} for available palettes. If omitted, the default palette is provided by `c4a_default_palette`. The palette name can be prefixed with a `"-"` symbol, which will reverse the palette (this can also be done with the `reverse` argument).
 #' @param n number of colors. If omitted then: for type `"cat"` the maximum number of colors is returned, for types `"seq"`, `"div"`, and `"cyc"`, 7 , 9, and 9 colors respectively.
-#' @param m number of rows in case type is `"bivs"`, `"bivc"`, `"bivd"` or  `"bivg"` (which stand for respectively sequential, categorical, diverging and desaturated (`g` for 'grayscale')).
-#' @param type type of color palette, in case `palette` is not specified: one of `"cat"` (categorical/qualitative palette), `"seq"` (sequential palette), `"div"` (diverging palette), `"cyc"` (cyclic palette), and `"bivs"`/`"bivc"`/`"bivd"`/`"bivg"` (bivariate: respectively seq-seq seq-cat, seq-div, and seq-desaturated).
+#' @param m number of rows in case type is bivariate, so one of `"bivs"`, `"bivc"`, `"bivd"` or  `"bivg"` (see \code{\link{c4a_types}} for descriptions)
[…]

### 2024-10-09 R/c4a_options.R

@@ -1,0 +1,79 @@
+#' Set cols4all options
+#'
+#' Get or set global options for c4a. Works similar as the base function `options`
+#'
+#' @param ... Use character values to retrieve options. To set options, either use named arguments (where the names refer to the options), a list that consists of those options.
+#'
+#' @details
+#'
+#' | **Option**        | **Description**   |
+#' | ------------- | ------------- |
+#' | defaults		|  Default palettes per type |
+#' | CBF_th		|  Parameters that label a palette as color blind friendly  |
+#' | CBVF_th		|  Parameters that label a palette as very color blind friendly  |
+#' | CBU_th		| Parameters that label a palette as color blind unfriendly |
+#' | CrangeFair		| Maximum chroma range for which a palette is considered harmonic |
+#' | CrangeUnfair		| Minimum chroma range for which a palette is considered disharmonic  |
+#' | LrangeFair		| Maximum luminance range for which a palette is considered harmonic |
+#' | LrangeUnfair		| Minimum luminance range for which a palette is considered disharmonic |
+#' | Cintense		| Chroma of colors that are considered intense |
+#' | Cpastel		| Chroma of colors that are considered 'pastel' |
+#' | HwidthDivRainbow		| A diverging palette is labeled as 'rainbow hue' if HwidthL or HwidthR are at least `HwidthDivRainbow` |
+#' | HwidthDivSingle		| A diverging palette is labeled as 'single hue' if HwidthL and HwidthR are at most `HwidthDivSingle` |
+#' | HwidthSeqRainbow | A sequential palette is labeled as 'rainbow hue' if Hwidth is at least `HwidthSeqRainbow` |
+#' | HwidthSeqSingle | A sequential palette is labeled as 'single hue' if Hwidth is at most `HwidthSeqSingle` |
+#' | naming_fun | Function that returns a distance matrix with the `naming_colors` (see examples) |
+#' | naming_fun_args | List of arguments for `naming_fun` |
+#' | naming_colors | Vector of prototype colors for the color names (see examples) |
+#' | naming_softmax | List of parameters for the softmax function applied to the distance matrix |
+
+#'
+#' @md
+#' @name c4a_options
+#' @rdname c4a_options
+#' @return A list of options
+#' @example ./examples/c4a_options.R
[…]

### 2024-10-09 R/c4a_palettes.R

@@ -2,7 +2,7 @@
 #'
 #' `c4a_palettes` lists all available cols4all color palettes. Palettes are organized by series. The available series are listed with `c4a_series`. Palettes are also organized per functional type, where we currently support: categorical `"cat"`, sequential `"seq"`, diverging `"div"`", cyclic `"cyc"`, and bivariate (seq x seq `"bivs"`, seq x cat `"bivc"`, seq x div `"bivd"`, seq x desaturated `"bivg"`) palette types. The function `c4a_types` lists all available types. The function `c4a_overview` gives an overview table of the number of palette per series and type. In an IDE with auto-completion (such as RStudio) it is possible to browse through the palette names with `.P` (using `$` like in lists).
 #'
-#' @param type type of color palette: one of `"all"` (all palettes), `"cat"` (categorical/qualitative palettes), `"seq"` (sequential palettes), `"div"` (diverging palettes), `"cyc"` (cyclic), and one of the bivariate palette types: `"bivs"`, `"bivc"`, `"bivd"`, `"bivg"`
+#' @param type type of color palette: one of `"all"` (all palettes), `"cat"`, `"seq"`, `"div"`, `"cyc"`, `"bivs"`, `"bivc"`, `"bivd"`, or `"bivg"`. See \code{\link{c4a_types}} for descriptions.
 #' @param series series to list the palettes from. Run `c4a_series` to see the options.
 #' @param full.names should full names, i.e. with the prefix "series."? By default `TRUE`.
 #' @param as.data.frame should `c4a_series` and `c4a_types` return the result as a data.frame, with description included as a column?
@@ -80,9 +80,10 @@ c4a_types = function(series = NULL, as.data.frame = TRUE) {
 
 #' @rdname c4a_palettes
 #' @param return.matrix should only a matrix be returned with numbers per palette and type? If `FALSE` a data.frame is returned with addional information
+#' @param zero.count.as.NA should zeros counted in the table be returned as 0 (`FALSE`, default) or as `NA` (`TRUE`)?
 #' @name c4a_series
 #' @export
-c4a_overview = function(return.matrix = FALSE) {
+c4a_overview = function(return.matrix = FALSE, zero.count.as.NA = FALSE) {
 	z = .C4A$z
 
 	if (is.null(z)) {
@@ -95,6 +96,7 @@ c4a_overview = function(return.matrix = FALSE) {
[…]

### 2024-10-09 R/c4a_plot_hues.R

@@ -141,7 +141,7 @@ c4a_plot_hues = function(pal, dark = FALSE, C_grey = 5, width = c("none", "total
 
 			grid::grid.polyline(x = grid::unit(coords[,1], "npc"),
 								y = grid::unit(coords[,2], "npc"), gp = grid::gpar(col = fc, lwd = 1.5))
-			grid::grid.text(paste0(HW, "°"),
+			grid::grid.text(paste0(HW, "\u00b0"),
 							x = grid::unit(tcoords[1], "npc"),
 							y = grid::unit(tcoords[2], "npc"),
 							just = "right",

### 2024-10-09 R/c4a_scores.R

@@ -1,0 +1,25 @@
+#' Get information from a cols4all palette
+#'
+#' Get information from a cols4all palette
+#'
+#' @param palette name of the palette
+#' @param type type of palettes (in case palette is not specified)
+#' @param series series name (in case palette is not specified)
+#' @param n number of colors
+#' @param no.match what happens is no match is found? Options: `"message"`: a message is thrown with suggestions, `"error"`: an error is thrown, `"null"`: `NULL` is returned
+#' @param verbose should messages be printed?
+#' @return list with the following items: name, series, fullname, type, palette (colors), na (color), nmax, and reverse. The latter is `TRUE` when there is a `"-"` prefix before the palette name.
+#' @export
+#' @example examples/c4a_scores.R
+c4a_scores = function(palette = NULL, type = NULL, series = NULL, n = NA, no.match = c("message", "error", "null"), verbose = TRUE) {
+	if (!is.null(palette)) {
+		z = get_zp(palette, n, no.match = no.match, verbose = verbose)
+	} else {
+		if (is.null(type)) stop("Please specify either palette or type (optionally in combination with series)")
+		if (is.na(n)) n = .C4A$ndef[[type]]
+		pals = c4a_palettes(type = type, series = series)
+		z = do.call(rbind, lapply(pals, get_zp, no.match = no.match, verbose = verbose))
+
+	}
+	show_attach_scores(z)
+}

### 2024-10-09 R/c4a_table.R

@@ -483,8 +483,11 @@ plot_table = function(p, text.format, text.col, include.na, cvd.sim, verbose) {
 #'
 #' Graphical user interface to analyse palettes. `c4a_table` shows a table that can be opened in the browser. `c4a_gui` is a graphical user interface (shiny app) around this table.
 #'
+#' See vignette how the properties are calculated. Parameters, such as threshold values which determined when palettes are classified as "colorblind-friendly", can be specified via \code{\link{c4a_options}}. Also the nameability score function (which is in development) can be specified there. See the examples of \code{\link{c4a_options}} for both use cases.
+#'
 #' @param type type of palette. Run \code{\link{c4a_types}} to see the implemented types and their description. For `c4a_gui` it only determines which type is shown initially.
 #' @param n,m `n` is the number of displayed colors. For bivariate palettes `"biv"`, `n` and `m` are the number of columns and rows respectively. If omitted: for `"cat"` the full palette is displayed, for `"seq"`, `"div"` and `"cyc"`, 7, 9, and 9 colors respectively, and for `"bivs"`/`"bivc"`/`"bivd"`/`"bivg"` 4 columns and rows. For `c4a_gui` it only determines which number of colors initially.
+#' @param continuous should the palettes as continuous instead of discrete. Only applicable for `"seq"`, `"div"`, and `"cyc"`.
 #' @param filters filters to be applied. A character vector with a subset from:`"nmax"` (only palettes where `n = nmax`, which is only applicable for categorical palettes),  `"cbf"` (colorblind-friendly), `"fair"` (fairness),`"naming"` (nameability),  `"crW"` (sufficient contrast ratio with white), and `"crB"` (sufficient contrast ratio with black). By default an empty vector, so no filters are applied.
 #' @param cvd.sim color vision deficiency simulation: one of `"none"`, `"deutan"`, `"protan"`, `"tritan"`
 #' @param sort column name to sort the data. The available column names depend on the arguments `type` and `show.scores`. They are listed in the warning message. Use a `"-"` prefix to reverse the order.
@@ -509,7 +512,7 @@ plot_table = function(p, text.format, text.col, include.na, cvd.sim, verbose) {
 #' @export
 #' @return An HMTL table (`kableExtra` object)
 #' @rdname c4a_gui
-#' @name c4a_gui
+#' @name c4a_table
[…]

### 2024-10-09 R/onLoad.R

@@ -35,80 +35,6 @@
 "_PACKAGE"
 
 
-#' Set cols4all options
-#'
-#' Get or set global options for c4a. Works similar as the base function `options`
-#'
-#' @param ... Use character values to retrieve options. To set options, either use named arguments (where the names refer to the options), a list that consists of those options.
-#'
-#' @details
-#'
-#' | **Option**        | **Description**   |
-#' | ------------- | ------------- |
-#' | defaults		|  Default palettes per type |
-#' | CBF_th		|  Parameters that label a palette as color blind friendly  |
-#' | CBU_th		| Parameters that label a palette as color blind unfriendly |
-#' | CrangeFair		| Maximum chroma range for which a palette is considered harmonic |
-#' | CrangeUnfair		| Minimum chroma range for which a palette is considered disharmonic  |
-#' | LrangeFair		| Maximum luminance range for which a palette is considered harmonic |
-#' | LrangeUnfair		| Minimum luminance range for which a palette is considered disharmonic |
-#' | Cintense		| Chroma of colors that are considered intense |
-#' | Cpastel		| Chroma of colors that are considered 'pastel' |
-#' | HwidthDivRainbow		| A diverging palette is labeled as 'rainbow hue' if HwidthL or HwidthR are at least `HwidthDivRainbow` |
-#' | HwidthDivSingle		| A diverging palette is labeled as 'single hue' if HwidthL and HwidthR are at most `HwidthDivSingle` |
-#' | HwidthSeqRainbow | A sequential palette is labeled as 'rainbow hue' if Hwidth is at least `HwidthSeqRainbow` |
-#' | HwidthSeqSingle | A sequential palette is labeled as 'single hue' if Hwidth is at most `HwidthSeqSingle` |
-#'
-#' @md
-#' @name c4a_options
-#' @rdname c4a_options
-#' @return A list of options
-#' @export
-c4a_options = function(...) {
-	lst = list(...)
-	e1 = parent.frame()
-	nms = c("defaults", "CBF_th", "CBU_th", "CrangeFair", "CrangeUnfair", "LrangeFair", "LrangeUnfair", "Cintense", "Cpastel", "HwidthDivRainbow", "HwidthDivSingle", "HwidthSeqRainbow", "HwidthSeqSingle", "naming_fun", "naming_fun_args", "naming_colors", "naming_softmax")
-
-	o = as.list(.C4A)[nms]
-
[…]

### 2024-10-09 examples/c4a.R

@@ -1,22 +1,38 @@
-c4a_palettes("div")
+# get the colors from brewer.set3 and plot them
+set3 <- c4a("brewer.set3")
+c4a_plot_hex(set3, nrows = 1)
 
-c4a(type = "cat")
+c4a("hcl.set2", n = 36) |> c4a_plot_hex()
+c4a("-hcl.set2", n = 12) |> c4a_plot_hex()
 
-(pal = c4a("tol.sunset", n = 7, range = c(0, .6)))
+# how to know which palettes are avaiable?
+# 1) Via the interactive tool:
+\dontrun{
+	c4a_gui()
+}
 
-c4a_plot(pal)
+# 2) Via the overview function:
+c4a_palettes(type = "cat")
+c4a_palettes(series = "brewer")
+c4a_palettes(type = "cat", series = "brewer")
+# Run c4a_overview() to see which are available
 
-c4a("set2")
+# 3) Via .P
+.P$brewer$cat$set3
 
-c4a("hcl.set2")
+# each palette contains a color for missing values
 
-c4a("hcl.set2", n = 8)
+c4a("carto.safe", 7)
+c4a_na("carto.safe")
 
-# reversed palette
-c4a("hcl.set2", reverse = TRUE, n = 8)
+c4a_plot_hex("carto.safe", n = 7, include.na = TRUE)
 
-# handy shortcut
-c4a("-hcl.set2", n = 8)
 
-# the color for missing values is white:
-c4a_na("hcl.set2")
+c4a_plot_hex("carto.safe", n = 7, include.na = TRUE)
+# same (but shorter) as
+# c4a_plot_hex(c(c4a("carto.safe", 7), c4a_na("carto.safe")), include.na = TRUE)
+
+
+# color ramp
+c4a("viridis", 100) |> c4a_plot()
+c4a_ramp("viridis")(100) |> c4a_plot()

### 2024-10-09 examples/c4a_options.R

@@ -1,0 +1,65 @@
+# Example how to lower the color-blind friendly threshold
+# for categorical palettes (so more smileys in the GUI!)
+# CBF_th: one smiley
+# CBVF_th: two smileys
+
+# current table
+\dontrun{
+c4a_table(n = 9, sort = "cbfriendly")
+
+opts = c4a_options("CBF_th", "CBVF_th")
+opts$CBF_th$cat["min_dist"] = 7
+opts$CBVF_th$cat["min_dist"] = 10
+
+
+old = c4a_options(opts)
+
+# more smileys :-) :-)
+c4a_table(n = 9, sort = "cbfriendly")
+
+# set the old settings back
+c4a_options(old)
+}
+
+# Example how to use own nameability function
+#
+# This function should:
+# - have an argument "pal" (vector of colors)
+# - optionally have other arguments
+# - return a distance matrix of n rows (length of pal) and k columns (classes).
+#   It shoud have columns names that correspond to the naming colors (see below).
+naming_RGB = function(pal) {
+	cols = colorspace::hex2RGB(pal)
+	coords = cols@coords
+
+	cls = apply(coords, MARGIN = 1, which.max)
+	mx = apply(coords, MARGIN = 1, max)
+	dominance = ((mx + 0.001) / (rowSums(coords) + 0.001))
+	cls[dominance < 0.4] = 4L
+
+	m = matrix(0, nrow = length(pal), ncol = 4,
+        dimnames = list(NULL, c("Red", "Green", "Blue", "Other")))
+	for (i in 1:nrow(m)) {
+		m[i, cls[i]] = 1
+	}
+
+	-m
+}
+
+# testing this function...
+naming_RGB(c4a("brewer.set1")) #fair enough
+
+# This vector should contain the 'prototype' colors, and have names that correspond
+# to the column names of the returned matrices by the function above.
+names_RGB =
+    c("Red" = "#FF0000",
+      "Green" = "#00FF00",
+      "Blue" = "#0000FF",
+      "Other" = "#AAAAAA")
+
+# Set the options (may take a while because if calculated the nameability scores)
+\dontrun{
+c4a_options(naming_fun = naming_RGB,
+			naming_fun_args = list(),
+			naming_colors = names_RGB)
+}

### 2024-10-09 examples/c4a_plot.R

@@ -1,12 +1,17 @@
 c4a_plot("brewer.set1", nrows=1)
 
-c4a_plot("greens", nrows=1)
+c4a_plot_hex("brewer.set1", nrows=1)
 
-c4a_plot("tol.pu_gn", nrows=1)
+c4a_plot_cvd("brewer.set1")
 
-c4a_plot(.P$c4a$bivs$pu_gn_bivs, n = 5)
+c4a_plot_cvd("greens")
 
-c4a_plot(.P$c4a$bivd$pu_gn_bivd, n = 5)
+c4a_plot_cvd("tol.pu_gn")
 
-c4a_plot(.P$c4a$bivg$gn_bivg, n = 5)
+c4a_plot(.P$cols4all$bivs$pu_gn_bivs, n = 5)
 
+c4a_plot(.P$met$bivc$monet)
+
+c4a_plot(.P$cols4all$bivd$pu_gn_bivd, n = 5)
+
+c4a_plot(.P$cols4all$bivg$gn_bivg, n = 5)

### 2024-10-09 README.Rmd
0.8 ready
vignette updated, figures working?

@@
 
-[<img src="vignettes/cols4all2.png" width="900"/>](vignettes/cols4all2.png)
+[<img src="vignettes/cols4all2.jpg" width="900"/>](vignettes/cols4all2.jpg)
 
@@
 
-[<img src="vignettes/cols4all_fair_sort_cbf.png" width="900"/>](vignettes/cols4all_fair_sort_cbf.png)
+[<img src="vignettes/cols4all_fair_sort_cbf.jpg" width="900"/>](vignettes/cols4all_fair_sort_cbf.jpg)
 
@@
 
-[<img src="vignettes/cols4all_div.png" width="900"/>](vignettes/cols4all_div.png)
+[<img src="vignettes/cols4all_div.jpg" width="900"/>](vignettes/cols4all_div.jpg)
 
@@
 
-[<img src="vignettes/c4a_pals.png" width="900"/>](vignettes/c4a_pals.png)
+[<img src="vignettes/c4a_pals.jpg" width="900"/>](vignettes/c4a_pals.jpg)
 
@@
 
-[<img src="vignettes/area7.png" width="900"/>](vignettes/area7.png)
+[<img src="vignettes/area7.jpg" width="900"/>](vignettes/area7.jpg)
 
@@
 
-[<img src="vignettes/area7d.png" width="900"/>](vignettes/area7d.png)
+[<img src="vignettes/area7d.jpg" width="900"/>](vignettes/area7d.jpg)
 
@@
 
-[<img src="vignettes/line7_wt.png" width="440"/>](vignettes/line7_wt.png)
-[<img src="vignettes/line7_bk.png" width="440"/>](vignettes/line7_bk.png)
+[<img src="vignettes/line7_wt.jpg" width="440"/>](vignettes/line7_wt.jpg)
+[<img src="vignettes/line7_bk.jpg" width="440"/>](vignettes/line7_bk.jpg)
 
@@
 
-[<img src="vignettes/friendly13.png" width="900"/>](vignettes/friendly13.png)
+[<img src="vignettes/friendly13.jpg" width="900"/>](vignettes/friendly13.jpg)
 

### 2024-10-16 README.Rmd
moved vignette to meet CRAN policy

@@
 Currently, this package contains palettes from several popular and lesser known color palette series.
-Own palettes series can be added as well.
+Users can also add their own palette series.
 
@@
 
-This gave us inspiration to develop own palettes: see these `cols4all` palettes below.
+This inspired us to develop our own palettes: see these `cols4all` palettes below.
 
@@
 
-Say we need a diverging palette that is color blind friendly, and what to choose one by eye. Then filter by "Colorblind-friendly" and sort by "Huddle Middle L" (the hue of the left wing):
+Say we need a diverging palette that is color blind friendly, and what to choose one by eye. Then filter by "Colorblind-friendly" and sort by "Hue Middle L" (the hue of the left wing):
 
@@
 
-We applied a basic heuristics to explore palettes that score well on a mix of the properties named above
+We applied a basic heuristic to explore palettes that score well on a mix of the properties named above
 
@@
 
-## Other R funtions
+## Other R functions
 
@@
 
-There are a few other pacakges with a large collection of color palettes, in particular [**pals**](https://kwstat.github.io/pals/) and [**paletteer**](https://github.com/EmilHvitfeldt/paletteer).
-There are a few features that distinguishes **cols4all** from those packages:
+There are a few other packages with a large collection of color palettes, in particular [**pals**](https://kwstat.github.io/pals/) and [**paletteer**](https://github.com/EmilHvitfeldt/paletteer).
+There are a few features that distinguish **cols4all** from those packages:
 

### 2024-10-16 README.Rmd
updated readme vignette images urls

@@
 
-This development version will be on CRAN soon (October 2024)
-
 ```{r, echo=FALSE, message=FALSE}
@@
 
-[<img src="vignettes/cols4all2.jpg" width="900"/>](vignettes/cols4all2.jpg)
+[<img src="vignettes/articles/cols4all2.jpg" width="900"/>](vignettes/articles/cols4all2.jpg)
 
@@
 
-[<img src="vignettes/cols4all_fair_sort_cbf.jpg" width="900"/>](vignettes/cols4all_fair_sort_cbf.jpg)
+[<img src="vignettes/articles/cols4all_fair_sort_cbf.jpg" width="900"/>](vignettes/articles/cols4all_fair_sort_cbf.jpg)
 
@@
 
-[<img src="vignettes/cols4all_div.jpg" width="900"/>](vignettes/cols4all_div.jpg)
+[<img src="vignettes/articles/cols4all_div.jpg" width="900"/>](vignettes/articles/cols4all_div.jpg)
 
@@
 
-[<img src="vignettes/c4a_pals.jpg" width="900"/>](vignettes/c4a_pals.jpg)
+[<img src="vignettes/articles/c4a_pals.jpg" width="900"/>](vignettes/articles/c4a_pals.jpg)
 
@@
 
-[<img src="vignettes/area7.jpg" width="900"/>](vignettes/area7.jpg)
+[<img src="vignettes/articles/area7.jpg" width="900"/>](vignettes/articles/area7.jpg)
 
@@
 
-[<img src="vignettes/area7d.jpg" width="900"/>](vignettes/area7d.jpg)
+[<img src="vignettes/articles/area7d.jpg" width="900"/>](vignettes/articles/area7d.jpg)
 
@@
 
-[<img src="vignettes/line7_wt.jpg" width="440"/>](vignettes/line7_wt.jpg)
-[<img src="vignettes/line7_bk.jpg" width="440"/>](vignettes/line7_bk.jpg)
+[<img src="vignettes/articles/line7_wt.jpg" width="440"/>](vignettes/articles/line7_wt.jpg)
+[<img src="vignettes/articles/line7_bk.jpg" width="440"/>](vignettes/articles/line7_bk.jpg)
 
@@
 
-[<img src="vignettes/friendly13.jpg" width="900"/>](vignettes/friendly13.jpg)
+[<img src="vignettes/articles/friendly13.jpg" width="900"/>](vignettes/articles/friendly13.jpg)
 

### 2024-10-16 README.Rmd
...

--- 2024-10-16 README.Rmd
@@
 
-[<img src="vignettes/cols4all2.jpg" width="900"/>](vignettes/cols4all2.jpg)
+[<img src="articles/cols4all2.jpg" width="900"/>](articles/cols4all2.jpg)
 
@@
 
-[<img src="vignettes/cols4all_fair_sort_cbf.jpg" width="900"/>](vignettes/cols4all_fair_sort_cbf.jpg)
+[<img src="articles/cols4all_fair_sort_cbf.jpg" width="900"/>](articles/cols4all_fair_sort_cbf.jpg)
 
@@
 
-[<img src="vignettes/cols4all_div.jpg" width="900"/>](vignettes/cols4all_div.jpg)
+[<img src="articles/cols4all_div.jpg" width="900"/>](articles/cols4all_div.jpg)
 
@@
 
-[<img src="vignettes/c4a_pals.jpg" width="900"/>](vignettes/c4a_pals.jpg)
+[<img src="articles/c4a_pals.jpg" width="900"/>](articles/c4a_pals.jpg)
 
@@
 
-[<img src="vignettes/area7.jpg" width="900"/>](vignettes/area7.jpg)
+[<img src="articles/area7.jpg" width="900"/>](articles/area7.jpg)
 
@@
 
-[<img src="vignettes/area7d.jpg" width="900"/>](vignettes/area7d.jpg)
+[<img src="articles/area7d.jpg" width="900"/>](articles/area7d.jpg)
 
@@
 
-[<img src="vignettes/line7_wt.jpg" width="440"/>](vignettes/line7_wt.jpg)
-[<img src="vignettes/line7_bk.jpg" width="440"/>](vignettes/line7_bk.jpg)
+[<img src="articles/line7_wt.jpg" width="440"/>](articles/line7_wt.jpg)
+[<img src="articles/line7_bk.jpg" width="440"/>](articles/line7_bk.jpg)
 
@@
 
-[<img src="vignettes/friendly13.jpg" width="900"/>](vignettes/friendly13.jpg)
+[<img src="articles/friendly13.jpg" width="900"/>](articles/friendly13.jpg)
 

### 2024-10-17 README.Rmd

@@
 
-[<img src="articles/cols4all2.jpg" width="900"/>](articles/cols4all2.jpg)
+[<img src="man/figures/cols4all2.jpg" width="900"/>](man/figures/cols4all2.jpg)
 
@@
 
-[<img src="articles/cols4all_fair_sort_cbf.jpg" width="900"/>](articles/cols4all_fair_sort_cbf.jpg)
+[<img src="man/figures/cols4all_fair_sort_cbf.jpg" width="900"/>](man/figures/cols4all_fair_sort_cbf.jpg)
 
@@
 
-[<img src="articles/cols4all_div.jpg" width="900"/>](articles/cols4all_div.jpg)
+[<img src="man/figures/cols4all_div.jpg" width="900"/>](man/figures/cols4all_div.jpg)
 
@@
 
-[<img src="articles/c4a_pals.jpg" width="900"/>](articles/c4a_pals.jpg)
+[<img src="man/figures/c4a_pals.jpg" width="900"/>](man/figures/c4a_pals.jpg)
 
@@
 So ideal for maps and other space-filling visualizations!
-These are used in [https://github.com/r-tmap/tmap](`tmap4`).
+These are used in [`tmap4`](https://github.com/r-tmap/tmap).
 
-[<img src="articles/area7.jpg" width="900"/>](articles/area7.jpg)
+[<img src="man/figures/area7.jpg" width="900"/>](man/figures/area7.jpg)
 
@@
 
-[<img src="articles/area7d.jpg" width="900"/>](articles/area7d.jpg)
+[<img src="man/figures/area7d.jpg" width="900"/>](man/figures/area7d.jpg)
 
@@
 
-[<img src="articles/line7_wt.jpg" width="440"/>](articles/line7_wt.jpg)
-[<img src="articles/line7_bk.jpg" width="440"/>](articles/line7_bk.jpg)
+[<img src="man/figures/line7_wt.jpg" width="440"/>](man/figures/line7_wt.jpg)
+[<img src="man/figures/line7_bk.jpg" width="440"/>](man/figures/line7_bk.jpg)
 
@@
 
-[<img src="articles/friendly13.jpg" width="900"/>](articles/friendly13.jpg)
+[<img src="man/figures/friendly13.jpg" width="900"/>](man/figures/friendly13.jpg)
 
